\documentclass[11pt]{article}

\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{newtxtext,newtxmath}
\usepackage[margin=1.05in]{geometry}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{graphicx}
\usepackage{microtype}
\usepackage{xcolor}
\usepackage[hidelinks]{hyperref}
\usepackage[nameinlink,noabbrev]{cleveref}

\hypersetup{
  pdftitle={Local Dephasing in a de Sitter Static Patch},
  pdfauthor={Brian Fioca},
  pdfsubject={The sharp localization-energy cost of a compact scalar pointer channel}
}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

\newcommand{\R}{\mathbb R}
\newcommand{\cD}{\mathcal D}
\newcommand{\cE}{\mathcal E}
\newcommand{\cH}{\mathcal H}
\newcommand{\dd}{\,\mathrm d}
\newcommand{\eobs}{\epsilon_{\mathrm{obs}}}
\newcommand{\EK}{E_{K}}
\newcommand{\Tr}{\operatorname{Tr}}
\newcommand{\supp}{\operatorname{supp}}
\newcommand{\sech}{\operatorname{sech}}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
\newcommand{\abs}[1]{\left\lvert#1\right\rvert}
\newcommand{\ip}[2]{\left\langle#1,#2\right\rangle}

\title{Local Dephasing in a de Sitter Static Patch:\\
Sharp Localization--Energy Cost of a Compact Pointer Channel}
\author{Brian Fioca}
\date{Draft of June 2026}

\begin{document}
\maketitle

\begin{abstract}
We consider a degenerate pointer qubit coupled through a smooth compact
spacetime source to the conformally coupled massless scalar in a
four-dimensional de Sitter static patch.  The exact Weyl-displacement channel
and its thermal dephasing factor are known.  We eliminate the source strength
and solve the resulting finite-support energy optimization.  If the source is
confined to areal radius $a<R$ for static time $T$, set
$y=\operatorname{arctanh}(a/R)+T/R$.  The halved diamond distance from
complete pointer dephasing obeys
\[
 \log\frac{1}{2\eobs}
 \leq \EK R C_{\rm opt}(y),
 \qquad C_{\rm opt}(y)=2y\Lambda(y/2)\leq F(y),
 \qquad
 F(y)=\frac{4\operatorname{arsinh}(1)}{\pi}y
      +\frac{8}{\pi^3}y^2 .
\]
Here $\Lambda(\tau)$ is the simple top eigenvalue of the positive compact
kernel $\pi^{-1}\log[\sinh(\tau(u+v))/\sinh(\tau|u-v|)]$ on $(0,1)$, and
$\EK$ is the exact post-switch scalar Killing-energy excess generated by the
same source.  The optimization includes arbitrary angular and canonical data;
its unique extremizing sector is s-wave momentum.  Temperature corrects the
small-support vacuum cost only at cubic order, while $8y^2/\pi^3$ is the sharp
leading large-support coefficient.  Smooth compact sources approach the
optimum when their spatial worldtube approaches the final support ball.  For
flux-free spherical final data, the same scalar datum gives exact
vanishing-extrinsic-curvature Einstein--scalar Hamiltonian and momentum
constraint data with mass $m=\EK$.  We also prove
that the dephasing functional is incomparable with the local coherent-state
entropy entering recent Klein--Gordon Bekenstein bounds.  The result concerns
field energy and final-slice constraints.  It does not include the material
stress of an autonomous switching device, derive the channel on the perturbed
geometry, or solve the coupled Einstein--matter evolution.
\end{abstract}

\input{sections/introduction}
\input{sections/model_and_channel}
\input{sections/localization_theorem}
\input{sections/achievability}
\input{sections/gravity}
\input{sections/comparison}
\input{sections/discussion}

\begingroup
\small
\let\originalthebibliography\thebibliography
\renewcommand{\thebibliography}[1]{%
  \originalthebibliography{#1}%
  \setlength{\itemsep}{0pt}%
  \setlength{\parskip}{0pt}%
}
\bibliographystyle{unsrt}
\bibliography{references}
\endgroup

\end{document}
