\section{Reproducibility and certificate provenance}
\label{sec:reproducibility}

The proof is split between human-checkable analytic reductions and a
substantial exact-rational validation code base.  The checker implements
rational interval arithmetic,
rational Taylor or Lagrange enclosures for the required transcendental
functions, interval automatic differentiation, fixed-grid outward rounding,
and exact comparison of the radii-polynomial inequalities.  It does not call a
floating transcendental library in a theorem comparison.

The two primary archives are listed in \cref{tab:archives}.  The hashes are
of the files as distributed with this manuscript's repository snapshot.

\begin{table}[ht]
  \centering
  \small
  \begin{tabular}{@{}p{0.31\linewidth}p{0.38\linewidth}p{0.22\linewidth}@{}}
    \toprule
    Result & Archive & SHA-256 prefix \\
    \midrule
    Profile and spectral endpoint data & combined AU.1/AU.2 certificate & \texttt{1d5fe53786cc} \\
    Fixed-wall radial gap & radial-gap certificate & \texttt{695310609d07} \\
    \bottomrule
  \end{tabular}
  \caption{Primary exact-rational proof archives.  Full hashes are recorded in
  the accompanying manifest and in the text below.}
  \label{tab:archives}
\end{table}

The full digests are
\begin{align*}
 &\texttt{1d5fe53786cc280006d7b1092d360556d4d8d8684e5ae3356ce8cd6d084e72a9},\\
 &\texttt{695310609d070f6dba2c678982ffc87472ed7c341813137fb90ad9dd9e1e6096}.
\end{align*}
The radial certificate additionally consumes the authenticated sharp-tube
snapshot with digest
\begin{center}
\texttt{1781a2ff357f3b165d23e290eb403552d77d099a72b9e90920735e6a80b30431}.
\end{center}

The union of application sources pinned by the two certificate ledgers and the
sharp-tube snapshot is 14,059 physical lines in this repository snapshot,
spread across fourteen modules and audit drivers.  It is a source-pinned
review surface, not a minimized or formally verified trusted kernel: it
includes proposal and serialization paths, while the interpreter and standard
library remain environmental assumptions.  The certificate records a SHA-256
digest for every file in this application surface.  Floating-point routines
may propose splines and meshes; exact-rational validation modules recompute
every proof-relevant enclosure and closing inequality.

\begin{center}
\begin{minipage}{0.96\linewidth}
  \small
  \centering
  \refstepcounter{table}
  \label{tab:validation-source-roles}
  \begin{tabular}{@{}p{0.27\linewidth}p{0.48\linewidth}r@{}}
    \toprule
    \raggedright Primary reviewer role
      & \raggedright Ledger-pinned source group
      & Physical lines \\
    \midrule
    \raggedright Candidate generation
      & \raggedright
        \path{qgtoy/skyrmion_global_bvp_certificate_generator.py}; proposal only
      & 1,303 \\
    \raggedright Exact validation and reconstruction
      & \raggedright nine modules: rational arithmetic/text, BVP/origin, spectral,
        radial-gap, and sharp-profile validation
      & 11,024 \\
    \raggedright Closing theorem comparison
      & \raggedright three-validator subset: spectral derivatives/ledger and radial gap
      & subset \\
    \raggedright Presentation and serialization
      & \raggedright decimal summaries and manuscript; no theorem authority or pinned
        plotting module
      & 0 additional \\
    \raggedright Authentication, audit, and replay
      & \raggedright four \path{experiments/skyrmion_*_audit.py} drivers
      & 1,732 \\
    \midrule
    \multicolumn{1}{l}{Ledger-pinned application surface}
      & \multicolumn{1}{l}{fourteen files}
      & 14,059 \\
    \bottomrule
  \end{tabular}
  \par\smallskip
  \raggedright
  Table~\thetable: Roles in the source-pinned review surface.  Roles can
  overlap: the closing-comparison row identifies a proof-critical subset of
  the exact validation row and is not counted twice.  Source hashes establish
  provenance, not correctness by themselves.
\end{minipage}
\end{center}

From the repository root, the focused validation path is
\begin{verbatim}
python -m pip install -e . 'pytest>=8,<10'
PYTHONPATH=. python experiments/skyrmion_newton_linearization_audit.py \
  --omega 3/4 --tube-radius 1/250 \
  --tube-trigonometric-terms 12 \
  --tube-rounding-denominator 1000000000000000000 \
  --spectral-trigonometric-terms 24 \
  --spectral-atanh-terms 80 --spectral-pi-terms 80 \
  --origin-kernel-terms 20 \
  --output /tmp/skyrmion_au2_replay_summary.json \
  --exact-certificate-output /tmp/skyrmion_au2_replay.json
python paper/validated_skyrmion_profile/audit_package.py \
  --replay-au2 /tmp/skyrmion_au2_replay.json
PYTHONPATH=. python experiments/skyrmion_au3b_sharp_snapshot_audit.py \
  --canonical-au2 \
    experiments/skyrmion_au2_global_tail_exact_certificate.json \
  --output /tmp/skyrmion_sharp_snapshot_replay.json
PYTHONPATH=. python experiments/skyrmion_full_radial_gap_audit.py \
  --output /tmp/skyrmion_radial_gap_replay.json
sha256sum \
  /tmp/skyrmion_sharp_snapshot_replay.json \
  /tmp/skyrmion_radial_gap_replay.json
python -m pytest -q \
  tests/test_validated_skyrmion_profile.py \
  tests/test_validated_skyrmion_bvp.py \
  tests/test_validated_skyrmion_spectral_derivatives.py \
  tests/test_validated_skyrmion_spectral_ledger.py \
  tests/test_static_patch_skyrmion_tail.py \
  tests/test_skyrmion_radial_dynamical_gap.py \
  tests/test_validated_skyrmion_radial_gap.py \
  tests/test_validated_rational_text.py \
  tests/test_validated_skyrmion_profile_manuscript.py
\end{verbatim}
The two hashes printed by \texttt{sha256sum} must be the sharp-snapshot and
radial values displayed above.  The combined AU.1/AU.2 archive also records the
literal command and a dependency ledger.  Its full-file hash is therefore
environment-sensitive.  A byte-identical clean replay used the archived output
paths, Python 3.12.13, NumPy 2.2.6, and no SciPy.

Across supported environments, the audit removes only the top-level provenance
fields \texttt{command} and \nolinkurl{dependency_versions}, then checks the
canonical mathematical-payload digest
\begin{center}
\texttt{f1703ab8f2b3f908a1609d03349c60becdb287225374c91c136fb275053d0f85}.
\end{center}
All exact inputs, source hashes, parameters, and theorem outputs remain inside
that comparison.
The source, certificate archives, manifest, and manuscript are distributed at
\begin{center}
\url{https://github.com/bfioca/physics-and-information-theory}.
\end{center}
The immutable external-review snapshot is the Git tag
\texttt{paper-a-review-v1}.  A later journal version should cite its release
tag or archival DOI; the file-level hashes above identify the present referee
snapshot independently of mutable branch names.
The exact test filenames and a machine-readable manifest are checked as part
of the release audit; no unrelated observer-algebra test is a premise of the
paper.

The archive format records both exact fractions and downward/upward decimal
summaries.  Decimal values printed in the manuscript are presentation layers,
not inputs to subsequent proof steps.  Every composed certificate checks the
identity of its upstream archive before importing a wall slope, inertia,
endpoint jet, or Newton radius.
