\section{Computer-assisted validation}
\label{sec:validation}

This section gives the analytic skeleton of the computer-assisted proof.  The
proof has two logically separate parts.  On the regular-origin patch we use a
uniform Volterra contraction, with the initial slope as a parameter.  On the
positive-radius interval we use an augmented Newton--Kantorovich argument for
the profile and that slope together.  Candidate splines may be generated
numerically, but every premise used below is recomputed with rational interval
arithmetic.  In particular, convergence of a floating-point shooting method
is not a premise of the theorem.

Throughout this section
\[
 a=\frac1{16},\qquad c=4,\qquad I=[a,c],\qquad
 N(x)=1-\frac{x^2}{400}.
\]
For convenient reference, write the profile equation in variational form as
\begin{align}
 \cG(F)
  &=-\bigl(\mathsf p(x,F)F'\bigr)'
    +\frac12\mathsf p_F(x,F)(F')^2+W_F(x,F),
    \label{eq:validation-profile-operator}\\
 \mathsf p(x,F)
  &=N(x)\bigl(x^2+8\sin^2F\bigr),\nonumber\\
 W(x,F)
  &=\sin^2F+\frac{2\sin^4F}{x^2}+x^2(1-\cos F).
 \nonumber
\end{align}
The apparent singularity at $x=0$ is handled before this formula is used on
$I$.

\subsection{The regular-origin chart}

Set
\[
 g(x)=\pi-F(x)=x u(t),\qquad t=x^2,\qquad
 p(t)=g'(x)=u(t)+2t u'(t).
\]
Thus $p(0)=u(0)=b$, where $b=-F'(0)>0$ is the shooting parameter, and
\begin{equation}
 u_p(t)=\int_0^1 p(ts^2)\,\dd s.
 \label{eq:validation-u-from-p}
\end{equation}
Introduce the entire even kernels
\[
 \operatorname{sinc}_1(z)=\frac{\sin z}{z},\qquad
 \operatorname{sinc}_2(z)=\frac{\sin(2z)}{2z},
\]
with their values at zero defined by continuity.  Direct substitution in the
Euler--Lagrange equation gives a nonsingular Volterra map.  One useful form is
\begin{align}
 A_p(t)&=u_p(t)\operatorname{sinc}_1(\sqrt t\,u_p(t)),\nonumber\\
 D_p(t)&=1+8A_p(t)^2,\nonumber\\
 Q_p(t)&=2u_p(t)\operatorname{sinc}_2(\sqrt t\,u_p(t))
 \bigl[1+4A_p(t)^2+4N(\sqrt t\,)p(t)^2\bigr]-tA_p(t),\nonumber\\
 (\mathcal V p)(t)&=
 \frac{1}{2tN(\sqrt t\,)D_p(t)}\int_0^t Q_p(s)\,\dd s,
 \label{eq:validation-volterra-map}
\end{align}
where the last expression is continued analytically to $t=0$.  Formula
\eqref{eq:validation-volterra-map} is especially useful for validation: all
trigonometric dependence occurs through entire power series, and integration
divides the coefficient of $t^j$ by the exact integer $2(j+1)$.

\begin{lemma}[Uniform regular-origin family]
\label{lem:validation-origin-volterra}
There is a common rational slope interval containing the certified center
\(
 \bar b=1462763601523/925827031517
\)
on which \eqref{eq:validation-volterra-map} has a unique fixed point in the
weighted balls
\begin{equation}
 p_b(t)=b-3\gamma(b)t+t^2r_b(t),
 \qquad \norm{r_b}_{\infty}\leq R,\qquad 0\leq t\leq a^2.
 \label{eq:validation-origin-ball}
\end{equation}
Here $\gamma(b)$ is the cubic coefficient obtained from the exact constant and
linear coefficient identities.  The fixed point is twice continuously
differentiable in $b$.  Consequently
\[
 \Phi(b)=F_b(a),\qquad \Gamma(b)=F_b'(a)
\]
and their first two derivatives are rigorously enclosed.  The same validation
is repeated on a wider slope interval containing the complete Newton slope
tube used below.
\end{lemma}

\begin{proof}
The checker translates \eqref{eq:validation-volterra-map} by the first two
exact Taylor coefficients and evaluates it on the weighted remainder ball in
\eqref{eq:validation-origin-ball}.  It verifies three strict inequalities: the
Volterra denominator stays positive, the image maps the ball into itself, and
the Lipschitz constant is smaller than one.  The parameterized contraction
theorem then gives the fixed-point family.  First- and second-order interval
automatic differentiation is applied to the same contraction equation, not
to finite differences of endpoint boxes.  At the narrow center family the
verified data are
\[
 b\in[1.579953,1.579954],\qquad R=\frac{13}{10},
 \qquad a=\frac1{16}.
\]
For the nonlinear tube the checker uses the larger remainder ball and slope
interval recorded by that tube and rechecks the same three inequalities.
The exact rational coefficient identities, power-series remainder bounds,
and the interval tables for both slope ranges are delegated to the origin
certificate artifact.  This lemma asserts uniqueness only in the displayed
weighted fibers; it does not assert unrestricted uniqueness of a singular
initial-value problem.
\end{proof}

For the center profile, a separately validated quintic origin expansion is
nested strictly inside the cubic fiber in
\eqref{eq:validation-origin-ball}.  Uniqueness in that fiber identifies the
two fixed points.  This sharper point expansion is used to seed the rational
positive-radius candidate, while the cubic family supplies the $b$-derivative
bounds required by the augmented proof.

\subsection{Rational center and augmented boundary map}

The candidate on $I$ is a globally $C^2$, piecewise-quintic spline $\widehat F$
on a 43-cell graded rational mesh.  Every node, coefficient, and join is a
rational number.  Numerical integration is used only to propose these data.
The checker verifies the joins and reevaluates all residuals from the archived
coefficients.

Let
\[
 \chi_a(x)=\frac{c-x}{c-a},\qquad
 \chi_c(x)=\frac{x-a}{c-a}.
\]
Rather than assume that the rational candidate has exact boundary values, set
\begin{equation}
 \bar u=\widehat F-\chi_a\widehat F(a)-\chi_c\widehat F(c),
 \qquad
 \bar F=\bar u+\chi_a\Phi(\bar b).
 \label{eq:validation-endpoint-lift}
\end{equation}
Then $\bar F(a)=\Phi(\bar b)$ and $\bar F(c)=0$ exactly.  The interval width in
$\Phi(\bar b)$ is retained in every subsequent coefficient and residual
calculation.

We regard the endpoint values as part of the chart and define
\begin{equation}
 \cN(F,b)=\bigl(\cG(F),F'(a)-\Gamma(b)\bigr),
 \quad F(a)=\Phi(b),\quad F(c)=0.
 \label{eq:validation-augmented-map}
\end{equation}
The verified nonlinear residual is evaluated in a cancellation-preserving
form: after multiplication by $x^2$, it is collected into exact rational
polynomial amplitudes multiplying
\[
 1,\quad \cos F,\quad \sin(2F),\quad \cos(2F),\quad \cos(4F).
\]
Each full expression is summed as a centered Taylor model before its range is
taken.  This step is essential near the origin, where separately ranging the
divergence terms loses several orders of magnitude.

\begin{lemma}[Authenticated center residual]
\label{lem:validation-center-residual}
The endpoint-corrected center \eqref{eq:validation-endpoint-lift} and the
scalar boundary residual in \eqref{eq:validation-augmented-map} admit exact
rational interval enclosures on all 43 cells.  The spline joins, endpoint
lifts, five-harmonic identity, and all Taylor remainders are checked
independently of the candidate generator.
\end{lemma}

\begin{proof}
This is a finite exact-arithmetic verification once the harmonic identity has
been obtained by symbolic differentiation.  The proof archive stores the
rational profile coefficients and the accepted aggregate outputs.  On replay,
the checker reconstructs every cellwise residual from those coefficients and
recomputes the scalar bound entering the final Newton defect; it does not
accept a decimal summary or a serialized cellwise table as proof input.
\end{proof}

\subsection{Dirichlet inverse and scalar Schur complement}

The reference inverse is built from the self-adjoint Dirichlet operator at the
raw rational spline,
\begin{equation}
 A_0h=-(P_0h')'+Q_0h,\qquad h(a)=h(c)=0,
 \label{eq:validation-jacobi}
\end{equation}
where $P_0=\mathsf p(x,\widehat F)$ and $Q_0$ is obtained by differentiating
\eqref{eq:validation-profile-operator}.  The fixed positive comparison
function
\[
 v_*(x)=\frac{8}{(x-33/16)^2+4}
\]
satisfies a certified Barta inequality $A_0v_*/v_*\geq\alpha>0$.  The
ground-state identity therefore yields
\begin{equation}
 \langle h,A_0h\rangle
 =\int_I P_0v_*^2\left[(h/v_*)'\right]^2\dd x
  +\int_I\frac{A_0v_*}{v_*}h^2\dd x,
 \label{eq:validation-ground-state}
\end{equation}
so $A_0$ is injective and obeys the maximum principle.

For sharp pointwise control, the checker does not use the coarse energy bound
alone.  Exact rational splines $\widehat u$ and $\widehat v$, vanishing at $a$
and $c$, respectively, define the factored Green parametrix
\begin{equation}
 \widehat G(x,s)=
 \frac{\widehat u(\min\{x,s\})\widehat v(\max\{x,s\})}{w_0},
 \label{eq:validation-green-parametrix}
\end{equation}
where $w_0$ is the certified oriented principal Wronskian.  The residuals
$A_0\widehat u$ and $A_0\widehat v$ control both the Wronskian drift and the
operator defect.

\begin{lemma}[Green resolvent bounds]
\label{lem:validation-green-resolvent}
The operator $A_0$ has a bounded inverse from $L^\infty(I)$ to
$W^{2,\infty}(I)\cap W^{1,\infty}_0(I)$.  The parametrix in
\eqref{eq:validation-green-parametrix} has certified defect
\(
 \varepsilon<0.089015
\), and the exact inverse obeys
\begin{equation}
 \norm{A_0^{-1}}_{L^\infty\to C^0}<0.228138,\qquad
 \norm{A_0^{-1}}_{L^\infty\to C^1}<1.43213,\qquad
 \norm{A_0^{-1}}_{L^\infty\to C^2}<59.3126.
 \label{eq:validation-green-bounds}
\end{equation}
The same construction yields a separate row bound for each source cell.
\end{lemma}

\begin{proof}
The checker verifies $P_0>0$ and the Barta inequality on an adaptively refined
rational partition; the archived exact lower bound satisfies
$\alpha>1.02359$.  Applying $A_0$ to the integral operator defined by
\eqref{eq:validation-green-parametrix} gives $I+E$ with
$\norm{E}<1$.  Hence
\[
 A_0^{-1}=\widehat T(I+E)^{-1},
\]
and a Neumann correction converts the exact row integrals of $\widehat G$,
$\partial_x\widehat G$, and the second-derivative jump term into
\eqref{eq:validation-green-bounds}.  The audit replay reconstructs the Barta
leaves, Wronskian bounds, fundamental-solution residuals, and row integrals
from the archived rational splines.  The archive records their accepted
aggregate bounds and source digests rather than serializing the transient
adaptive tables.
\end{proof}

It remains to invert the one-dimensional shooting block.  Define
\[
 \tau(f)=(A_0^{-1}f)'(a).
\]
If $\kappa$ solves $A_0\kappa=0$, $\kappa(a)=1$, $\kappa(c)=0$, Green's
identity gives the exact representation
\begin{equation}
 \tau(f)=\frac1{P_0(a)}\int_I\kappa(x)f(x)\,\dd x.
 \label{eq:validation-trace-representer}
\end{equation}
An independently validated rational approximation to $\kappa$ supplies the
trace bound; it is not inferred from the Schur solution it will later test.

Let $H$ solve
\begin{equation}
 A_0H=0,\qquad H(a)=\Phi'(\bar b),\qquad H(c)=0,\qquad
 S=H'(a)-\Gamma'(\bar b).
 \label{eq:validation-schur}
\end{equation}
Use local coordinates
\begin{equation}
 F=\bar F+w+\beta H
 +\chi_a\left[\Phi(\bar b+\beta)-\Phi(\bar b)-\Phi'(\bar b)\beta\right],
 \label{eq:validation-augmented-chart}
\end{equation}
with $w(a)=w(c)=0$.  Both value boundary conditions are then exact for every
$(w,\beta)$.

\begin{lemma}[Augmented inverse]
\label{lem:validation-schur-inverse}
The scalar Schur factor in \eqref{eq:validation-schur} is nonzero; the final
audit gives $S>2.8918$.  For the reference block built from $A_0$, the inverse
has the form
\begin{equation}
 w=A_0^{-1}r,\qquad
 \beta=\frac{s-\tau(r)}{S},
 \label{eq:validation-block-inverse}
\end{equation}
for an interior residual $r$ and derivative-boundary residual $s$.
\end{lemma}

\begin{proof}
The candidate $H$ is assembled from a rational regular-sensitivity spline and
a rational fundamental solution so that its right endpoint vanishes exactly.
The checker subtracts the image of its full residual using the independently
certified trace representer \eqref{eq:validation-trace-representer}.  The
corrected interval remains strictly positive.  Since
\cref{lem:validation-green-resolvent} makes $A_0$ invertible, elementary block
elimination gives \eqref{eq:validation-block-inverse}.  The replay recomputes
the residual, $L^1$ representer, trace correction, and Schur enclosures before
accepting the aggregate output.  Keeping the representer calculation separate
is important: using $H$ itself to certify the norm of $\tau$ would be
circular.
\end{proof}

We equip the Dirichlet component with the graph norm
\[
 \norm{w}_{A_0}=\norm{A_0w}_{\infty}
\]
and the augmented coordinates with
\begin{equation}
 \norm{(w,\beta)}_{\cX}
   =\max\{\norm{w}_{A_0},\omega\abs{\beta}\},
 \qquad \omega=\frac34.
 \label{eq:validation-product-norm}
\end{equation}
The $C^j$ estimates in \eqref{eq:validation-green-bounds}, together with the
corresponding corrected bounds for $H$, make this graph ball a pointwise
$C^1$ tube.  This is the bridge needed for the sign claims below; an energy
ball alone would not control $F'$ pointwise.

\subsection{Reduced Hessian and the radii polynomial}

A direct interval bound for $D^2\cG$ is valid but too coarse because it charges
the global $C^2$ inverse norm in every cell.  The closing estimate uses the
equation satisfied by each first variation.  If
\[
 A_0h=g_h,\qquad A_0k=g_k,
\]
put
\begin{align*}
 m&=\mathsf p_{FF}F'+\mathsf p_{xF},\\
 z&=W_{FFF}-\mathsf p_{FF}F''-\mathsf p_{xFF}F'
       -\frac12\mathsf p_{FFF}(F')^2.
\end{align*}
In the following identity, $P_0,Q_0$, and $P_0'$ are the coefficients of the
fixed reference operator at $\widehat F$, whereas
$\mathsf p_F,m,z$ are evaluated at the profile $F$ in the Newton tube.

\begin{lemma}[Center-equation Hessian identity]
\label{lem:validation-reduced-hessian}
For twice differentiable first variations $h,k$,
\begin{align}
 D^2\cG(F)[h,k]
  ={}&\frac{\mathsf p_F}{P_0}(k g_h+h g_k)-\mathsf p_Fh'k'
  +\left(\frac{\mathsf p_FP_0'}{P_0}-m\right)(hk'+kh')
 \nonumber\\
 &+\left(z-2\frac{\mathsf p_FQ_0}{P_0}\right)hk.
 \label{eq:validation-reduced-hessian}
\end{align}
Consequently the nonlinear Hessian can be bounded cell by cell using only
the local $C^0$ and $C^1$ Green rows and the known sources $g_h,g_k$.
\end{lemma}

\begin{proof}
Expand \eqref{eq:validation-profile-operator} twice.  In the resulting
symmetric expression replace both second derivatives using
\[
 h''=\frac{Q_0h-P_0'h'-g_h}{P_0},\qquad
 k''=\frac{Q_0k-P_0'k'-g_k}{P_0}.
\]
Collecting terms gives \eqref{eq:validation-reduced-hessian}; the factor $2$
in the final coefficient comes from the two substitutions.  This is an exact
algebraic identity, not an integration-by-parts estimate.  The checker bounds
both the direct and reduced formulas and takes the smaller certified bound on
each cell.  The symbolic expansion is human-checkable; the checker recomputes
the coefficient ranges and selected cellwise bounds during the audit replay.
\end{proof}

The scalar component also retains cell localization.  If
$K_J=\int_J|\widehat\kappa|\,\dd x$ on a source cell $J$ and
\(
 \epsilon_\kappa=C_0\norm{A_0\widehat\kappa}_\infty
\), then nonnegative cell bounds $b_J$ obey
\begin{equation}
 T(b)=\frac1{P_0(a)}
  \sum_J\bigl(K_J+\epsilon_\kappa |J|\bigr)b_J.
 \label{eq:validation-weighted-trace}
\end{equation}
Thus a large near-origin Hessian value is not multiplied by a global trace
constant over the whole domain.

Let $B$ denote the exact bounded inverse of the reference augmented block
defined by $A_0$, $\tau$, and $S$.  Its norm bounds are certified through the
rational Green and Schur parametrices above.  Using the norm
\eqref{eq:validation-product-norm}, the checker establishes
\begin{align}
 Y&\geq\norm{B\cN(\bar F,\bar b)}_{\cX},\nonumber\\
 Z_0&\geq\norm{I-BD\cN(\bar F,\bar b)}_{\cX\to\cX},\nonumber\\
 \norm{B[D\cN(X)-D\cN(\bar X)]}_{\cX\to\cX}
   &\leq Z_2\norm{X-\bar X}_{\cX}
   \quad\text{for }\norm{X-\bar X}_{\cX}\leq r.
 \label{eq:validation-yz-bounds}
\end{align}
The $Z_0$ calculation includes the coefficient mismatch between the
$A_0$-based reference inverse and the derivative at the endpoint-corrected
center.  The affine origin lift in
\eqref{eq:validation-augmented-chart}, including the certified bounds on
$\Phi''$ and $\Gamma''$, is included in $Z_2$.

\begin{lemma}[Newton--Kantorovich closure]
\label{lem:validation-newton-closure}
For $\omega=3/4$ and $r=1/250$, exact rational comparison gives the
conservative bounds
\begin{equation}
 Y\leq0.002296,\qquad Z_0\leq0.000078,\qquad Z_2\leq78.733.
 \label{eq:validation-final-yz}
\end{equation}
The radii polynomial and contraction factor satisfy
\begin{equation}
 \rho(r)=Y+(Z_0-1)r+\frac12Z_2r^2<-0.00107,\qquad
 q(r)=Z_0+Z_2r<0.31501.
 \label{eq:validation-radii-polynomial}
\end{equation}
Hence \eqref{eq:validation-augmented-map} has a unique zero in the closed
$\cX$-ball of radius $1/250$ about the certified center.
\end{lemma}

\begin{proof}
Equations \eqref{eq:validation-yz-bounds} imply that
\[
 \mathcal T(X)=X-B\cN(X)
\]
maps the closed radius-$r$ ball into itself when $\rho(r)<0$, and has Lipschitz
constant at most $q(r)$.  The strict inequalities in
\eqref{eq:validation-radii-polynomial} therefore give a contraction.  The
numbers in \eqref{eq:validation-final-yz} are outward decimal summaries of
exact fractions; the checker compares the fractions themselves.  On replay it
reconstructs the 43-cell Hessian bounds, local Green rows, weighted-trace
contributions, affine-lift bounds, and operator-mismatch bounds.  The exact
certificate stores the accepted fractions and all rational spline inputs
needed for that replay, but not the transient adaptive tables.
\end{proof}

The uniqueness in \cref{lem:validation-newton-closure} is local and
coordinate-explicit: it is uniqueness in the augmented graph ball, not global
uniqueness among all finite-energy hedgehog profiles.

\subsection{Post-Newton physical checks}

The contraction theorem is complete before any physical sign is asserted.
The checker then propagates the same Newton ball to pointwise jets.  On the
origin patch, the Volterra enclosure gives $p(t)>0$, hence $F'(x)<0$.  On each
positive-radius cell, the local Green bounds, the corrected auxiliary, and the
quadratic origin lift produce a tube enclosure for $F'$.  Every such enclosure
has a strictly negative upper endpoint.  The final cell gives the certified
wall interval
\begin{equation}
 F'(4)\in[-0.09465,-0.08746].
 \label{eq:validation-wall-slope}
\end{equation}
No sampled-grid interpolation is used in either sign test.

Finally, the dimensionless rotor inertia is the integral of the nonnegative
density
\begin{equation}
 j(x,F,F')=\frac{2\pi}{3}\left[
 \frac{x^2\sin^2F}{N(x)}+4x^2\sin^2F(F')^2
 +\frac{4\sin^4F}{N(x)}\right].
 \label{eq:validation-inertia-density}
\end{equation}
The checker evaluates \eqref{eq:validation-inertia-density} by directed
interval quadrature on the authenticated positive-radius cells.  The origin
Volterra family supplies a finite analytic upper bound on $[0,a]$; the lower
bound there may be discarded because the density is nonnegative.  A strictly
positive lower sum on $[a,c]$ and the finite origin upper bound give
\begin{equation}
 21.149\leq \mathcal I_{\rm rot}\leq48.921.
 \label{eq:validation-inertia-bound}
\end{equation}
Thus the validated solution is strictly decreasing, has a strictly negative
wall slope, and has finite positive rotor inertia.

\subsection{Exact arithmetic and proof provenance}

The trusted numerical layer uses rational endpoints throughout.  It evaluates
$\pi$ with a rational Machin formula, encloses sine and cosine with rational
Taylor--Lagrange remainders, and carries interval automatic derivatives
algebraically.  For the final tube, selected nonnegative upper bounds are
rounded upward to denominator $10^{18}$, while interval lower and upper
endpoints are rounded outward.  Twelve-term trigonometric enclosures are used
in the tube calculation.  These rounding operations control denominator
growth; they do not insert binary floating-point values into a theorem
comparison.

The complete accepted input and output used by this paper are stored in
\begin{center}
\path{experiments/skyrmion_au2_global_tail_exact_certificate.json}.
\end{center}
The archive contains the 43 rational cells for the profile, Schur auxiliary,
left Green factor, and right Green/trace representer; exact output fractions;
the command line; dependency versions; and SHA-256 digests of the checker
sources.  Its
SHA-256 digest is
\begin{center}
\texttt{1d5fe53786cc280006d7b1092d360556d4d8d8684e5ae3356ce8cd6d084e72a9}.
\end{center}
The profile component also has a compact human-readable audit:
\begin{center}
\path{experiments/skyrmion_newton_reduced_hessian_rounded_audit_result.json}.
\end{center}
The combined archive binds those outputs to the current source snapshot and
the AU.2 endpoint calculation.
The adaptive interval tables used in
\cref{lem:validation-origin-volterra,lem:validation-center-residual,%
lem:validation-green-resolvent,lem:validation-schur-inverse,%
lem:validation-reduced-hessian,lem:validation-newton-closure}
are deterministically recomputed by the pinned checker.  The archive retains
the exact inputs and accepted aggregate outputs, and the body records the
analytic identities and final outward enclosures needed to audit their role.
Tamper-rejection tests separately perturb endpoints,
splines, operator parameters, provenance links, and closing inequalities and
require the checker to reject the altered input.
