\section{Relation to prior bounds and detector work}
\label{sec:comparison}

\subsection{Bekenstein entropy is a different quadratic form}

For a Klein--Gordon wave packet localized in a spatial region of half-width
$a$, Hollands, Longo, and Morsella prove the Bekenstein-type inequality
\begin{equation}
 S(\Phi\mid B)\leq2\pi a E,
 \label{eq:bekenstein-bound}
\end{equation}
where $S(\Phi\mid B)$ is the entropy quadratic form of the local standard
subspace~\cite{HollandsLongoMorsella2026}.  At second quantization this is the
relative entropy of the corresponding coherent excitation.  This powerful
result is directly adjacent to \cref{thm:thermal-half-line}, but it controls a
different functional.

We make the distinction explicit in the simplest common setting.  Consider
the massless conformal vacuum in four-dimensional Minkowski space, restrict to
radial $q=0$ data in a ball of radius $a$, and normalize $\norm{p}_2=1$.  The
conformal-ball modular Hamiltonian and the half-line dephasing kernel give
\begin{align}
 S_B[p]
 &=\frac{\pi}{2a}\int_0^a(a^2-x^2)p(x)^2\dd x,
 \label{eq:SB-form}\\
 \Gamma_0[p]
 &=\ip p{h_0^{-1}p}
 =\frac1\pi\int_0^a\!\int_0^a p(x)p(y)
 \log\frac{x+y}{\abs{x-y}}\dd x\dd y.
 \label{eq:Gamma0-form}
\end{align}

\begin{proposition}[Quadratic-form separation]
\label{prop:bekenstein-separation}
At fixed $a$, there is no finite constant $c$ such that
$\Gamma_0[p]\leq cS_B[p]$ for all smooth normalized radial momenta, and no
finite constant $c'$ such that $S_B[p]\leq c'\Gamma_0[p]$ for all such data.
\end{proposition}

\begin{proof}
Let $f\in C_c^\infty(0,1)$ be positive and normalized in $L^2(0,1)$.  For a
packet in a boundary layer, set
\begin{equation}
 p_\delta^{\rm wall}(x)
 =\delta^{-1/2}f\!\left(\frac{a-x}{\delta}\right).
\end{equation}
Substitution in \cref{eq:SB-form} gives
\begin{equation}
 S_B[p_\delta^{\rm wall}]
 =\pi\delta\int_0^1u f(u)^2\dd u+O(\delta^2).
 \label{eq:SB-wall}
\end{equation}
For the logarithmic kernel, scaling both variables gives
\begin{equation}
 \Gamma_0[p_\delta^{\rm wall}]
 =\frac{\delta}{\pi}\left(\int_0^1f(u)\dd u\right)^2
 \log\frac1\delta+O(\delta).
 \label{eq:Gamma-wall}
\end{equation}
Thus $\Gamma_0/S_B\to\infty$.

For a packet near the center, set
\begin{equation}
 p_\delta^{\rm ctr}(x)=\delta^{-1/2}f(x/\delta).
\end{equation}
Then
\begin{equation}
 S_B[p_\delta^{\rm ctr}]=\frac{\pi a}{2}+O(\delta^2),
 \qquad
 \Gamma_0[p_\delta^{\rm ctr}]=O(\delta),
 \label{eq:center-asymptotics}
\end{equation}
because the logarithmic kernel becomes scale invariant after
$x=\delta u$, $y=\delta v$, while the two integration measures leave one
factor of $\delta$.  Hence $S_B/\Gamma_0\to\infty$.
\end{proof}

The separation has a simple physical meaning.  The Bekenstein entropy is a
local modular-information functional whose conformal weight vanishes at the
ball boundary.  The zero-temperature dephasing exponent is a global
one-particle norm, equivalently an entangling-particle count for the exact
detector model.  At finite temperature, the factor $\coth(\beta h/2)$ enhances
infrared fluctuations further.  Thermal dephasing can increase even while the
distinguishability of two displaced thermal states decreases.  The two notions
of information should not be identified.

\subsection{Relation to DSW and exact detector work}

DSW decoherence describes the loss of coherence of superposed charges or
masses near Killing horizons through soft field modes
\cite{DSW2023,DSW2025}.  Gralla and Wei gave a general horizon formulation and
precise rates in electromagnetic and Klein--Gordon analogues
\cite{GrallaWei2024}; Li evaluated the local scalar, electromagnetic, and
gravitational cases in de Sitter~\cite{Li2025}.  These works establish the
physical importance of the low-frequency KMS sector that appears in
\cref{eq:gamma-sector}.

Our setup differs in its question.  We do not derive a new horizon decoherence
rate.  We start from the exact pointer channel and maximize its dephasing
functional subject to finite final support and post-switch field energy.  The
result is consequently a constraint on a local implementation, not a claim
that horizon modes require a large Killing energy.  Indeed, sufficiently soft
horizon radiation can carry negligible Killing energy when its support and
duration are allowed to grow.  In \cref{eq:source-bound}, that escape route is
visible in the growth of $C_{\rm opt}$ with $L=\ell+T$.

Barcellos and Landulfo already separate field background energy, individual
switching work, and communication work for the exact gapless-detector channel
\cite{BarcellosLandulfo2021}.  We use their same exact channel class.  The new
step is to eliminate the source amplitude, identify the exact compact KMS
kernel, and solve the full final-support optimization.  We therefore make no
novelty claim for Magnus termination, the KMS characteristic function, the
coherent stress identity, or switching-energy accounting.

Related Unruh--DeWitt-gate work represents relativistic communication
channels as bosonic dephasing channels whose parameters depend on coupling,
smearing, and switching data~\cite{AsplingLawler2023}.  That work concerns
channel structure and capacity rather than the post-switch field-energy and
final-support variational problem posed here.

Likewise, spectral extremization of Riesz-potential operators is not new
\cite{RuzhanskySuragan2016}, and largest-eigenvalue questions for logarithmic
potentials have an established literature~\cite{AnoopJohnson2025,
JohnsonVerma2026}.  There are two exact relationships that sharpen this
comparison.  First, let $\widetilde p$ be the odd extension of $p$ from
$(0,1)$ to $(-1,1)$.  The zero-temperature operator is precisely the odd
sector of the finite-interval logarithmic convolution:
\begin{equation}
 \frac1\pi\int_{-1}^{1}\log\frac1{\abs{u-v}}\,
 \widetilde p(v)\dd v
 =\int_0^1 k_0(u,v)p(v)\dd v,
 \qquad 0<u<1.
 \label{eq:odd-log-convolution}
\end{equation}
Polosin studies high-index eigenvalue and eigenfunction asymptotics of that
convolution in its even and odd sectors~\cite{Polosin2022}.  Thus the vacuum
convolution problem itself is not new; that work does not give the thermal
principal-support law or the uniform temperature remainders used here.

Second, for $\tau>0$, set $r=e^{-2\tau u}$ and
$s=e^{-2\tau v}$.  Directly,
\begin{equation}
 k_\tau(u,v)=\frac1\pi\log\frac{1-rs}{\abs{r-s}}.
 \label{eq:hyperbolic-cross-ratio}
\end{equation}
The ratio on the right is the inverse pseudo-hyperbolic distance between two
points on a diameter of the Poincare disk.  The change of variables turns
$L^2(0,1)$ into
$L^2((e^{-2\tau},1),\dd r/(2\tau r))$.  Hence our operator is a weighted
one-dimensional geodesic compression of the hyperbolic logarithmic Green
kernel, not the two-dimensional hyperbolic-area operator studied in
\cite{JohnsonVerma2026}.

At the quadratic-form level, odd extension also writes the localized operator
$B_{\beta,L}$ as a finite-interval Wiener--Hopf term minus its Hankel image,
generated by the singular even multiplier
$\abs{k}^{-1}\coth(\beta\abs{k}/2)$.  The zero-frequency singularities of the
two image terms cancel in their Dirichlet difference.  This places the
operator near work on singular truncated Wiener--Hopf inverses and
determinants~\cite{Widom2006,Kozlowski2008}, although those results concern
different spectral quantities and symbol hypotheses.

Thus the claim here is not a first eigenvalue theorem for an unspecified
positive compact kernel.  It is the identification of the exact reflected
thermal operator governing this local pointer channel, proof that it controls
the complete canonical and angular phase space, and extraction of uniform
support laws with explicit constants.  The comparisons above do not by
themselves establish priority; they make the remaining operator-theory
novelty question precise.
