\section{Discussion and claim boundary}
\label{sec:discussion}

The result has two layers.  The analytic engine is the exact localization
coefficient for the thermal inverse Hamiltonian.  The physical composition is
the finite-pointer purity bound and its insertion into a simple observer code.
Together they say that a field record of large second R\'enyi entropy cannot be
produced with simultaneously fixed support and bounded post-switch scalar
energy in this model.  The statement is quantitative and nonperturbative in
the source amplitude.

Compact final support is the decisive hypothesis.  Without it, arbitrarily
soft thermal modes prevent an energy norm from controlling the dephasing
form.  With it, the compressed inverse Hamiltonian is bounded.  At arbitrary
inverse temperature its norm is governed by
$\mathcal K_{\pi L/\beta}$; in conformal de Sitter, angular resolvent order and
a separate field-coordinate estimate show that the same s-wave momentum mode
controls the complete canonical and angular phase space.  The
temperature--curvature relation enters only in this all-sector reduction.

\subsection{What is established}

The exact statements are:
\begin{enumerate}[leftmargin=*,itemsep=2pt]
 \item a controlled-source realization of the finite-pointer Schur channel,
 with pairwise exponents fixed by conditional field differences;
 \item the arbitrary-$\beta$ coefficient
 $\mathcal C_\beta(L)=2L\Lambda(\pi L/\beta)$, its simple top mode, global
 bracket, and uniform support asymptotics;
 \item the conformal de Sitter reduction of every angular and canonical sector
 to that unique s-wave momentum optimizer;
 \item the purity and R\'enyi bounds
 \cref{eq:finite-pointer-purity-bound,eq:simple-entropy-bound}, with exact
 saturation for the symmetric binary top-mode record;
 \item the physical-purity insertion
 \cref{eq:harlow-physical-floor} into the Harlow--Usatyuk--Zhao simple random
 code for an orthogonal CRT-real matter pair; and
 \item the branchwise final-slice constraint estimate
 \cref{eq:branchwise-gravity-renyi} for spherical $q_i=0$ records.
\end{enumerate}

The finite-pointer theorem is not claimed to be globally sharp at arbitrary
dimension.  Simultaneous equality would require every nonzero pairwise
difference to align with one simple top eigenspace and all off-diagonal
exponents to coincide.  The binary construction meets those conditions; a
generic larger alphabet does not.

Sharpness of the operator coefficient refers to final Cauchy data supported
in $[0,L]$.  For a source cylinder with optical radius $\ell<L$ and duration
$T=L-\ell$, finite propagation proves the same upper bound but not
near-controllability of its top eigenfunction.  Smooth compact sources do
approach every final-support Rayleigh quotient when their worldtube approaches
the declared final-support ball.

\subsection{Observer-code interpretation}

The code insertion supplies a concrete bridge that the earlier binary
diamond-error comparison did not.  The conditional field record purifies the
physical pointer state, so its record-system purity is exactly
$\Tr\rho_P^2$.  For the specified orthogonal CRT-real pair, this is precisely
the observer factor left in the Haar second moment of
\cite{HarlowUsatyukZhao2025}.  The field theorem therefore converts a
localization and energy budget into a lower floor on that ensemble-averaged
squared inner-product fluctuation.

This remains narrower than a gravitational observer theorem.  The result
does not bound the error of every fixed encoding map, establish a
high-probability statement, or control all matter pairs at once.  It also
does not identify the scalar energy with a total observer mass.  Harlow's
Hartle--Hawking extension separately argues that the formal
$S_{\rm Ob}\to\infty$ limit likely requires an infinitely heavy
observer~\cite{Harlow2026Observers}.  Our result realizes a related
energy--entropy obstruction in one field-record sector; it is not a
derivation of that broader claim.

The field-energy quantity is the post-switch scalar Killing energy of the
coherent conditional displacements.  It excludes the work, clock, battery,
trap, material support, and stress needed to implement the prescribed
sources.  The shared finite pointer is a gapless-detector idealization rather
than an autonomous relativistic probe field.  Accordingly, the theorem is not
a universal measurement-energy principle.

The gravity corollary is likewise deliberately branchwise and kinematical.
Each conditional spherical datum is placed on its own exact
Einstein--scalar final slice and must satisfy the local budget separately.
The quantum channel is still evaluated on the unperturbed de Sitter
background.  We do not derive a common branch-dependent quantum evolution,
include actuator stress, or prove global closeness to pure de Sitter; the
compact mass shifts the cosmological horizon.

\subsection{Novelty boundary}

Exact gapless-detector channels, their thermal damping factors, and their
field-energy accounting are prior art.  Positive compact kernels, simple top
eigenvalues, logarithmic-potential operators, and Riesz-potential
variational principles are also established mathematics.  The candidate
contribution is their specific composition here: the reflected thermal
operator selected by final support, its exact all-sector de Sitter norm, the
finite-pointer R\'enyi resource bound, and the resulting observer-code
mean-square floor.

The localized Klein--Gordon entropy comparison in \cref{sec:comparison}
shows that this purity functional is not a repackaging of the recent
Bekenstein bound.  It is not a second headline.  Independent review in
relativistic detector theory, integral-operator or fractional-Sobolev theory,
and quantum gravity remains the binding publication gate.

\subsection{Extensions}

A first physical strengthening would replace the Haar mean-square statement
by a concentration or high-probability theorem over the code ensemble.  That
requires controlling correlations among matter pairs and is not supplied by
the present second moment.  A second would replace the prescribed pointer by
a localized relativistic probe field, following the detector-from-field
framework of \cite{PercheEtAl2024}, while retaining the localization theorem
and accounting for the probe stress tensor.

The direct mathematical extension is the true fixed-source-cylinder
optimization.  It requires either a near-controllability theorem for sources
in a prescribed smaller worldtube or a source-to-final-data singular-value
problem replacing $\Lambda$.  A broader geometric extension would identify
conditions under which localized norms of
$h^{-1}\coth(\beta h/2)$ admit comparable sharp constants on other static
spacetimes.

\subsection{Reproducibility}

From the repository root, the analytic records, independent finite-pointer
replay, numerical illustration, and focused checks are reproduced by
\begin{verbatim}
PYTHONPATH=. python experiments/local_scalar_observer_cost_audit.py
PYTHONPATH=. python experiments/local_scalar_observer_spectrum.py
PYTHONPATH=. python experiments/finite_pointer_observer_audit.py
PYTHONPATH=. python experiments/finite_pointer_observer_clean_room_check.py
PYTHONPATH=. python -m pytest -q \
  tests/test_local_scalar_observer_cost.py \
  tests/test_local_scalar_observer_manuscript.py \
  tests/test_local_scalar_observer_spectrum.py \
  tests/test_finite_pointer_observer.py
\end{verbatim}
The frozen records check formulas, source hashes, numerical convergence, and
artifact provenance.  They do not replace the analytic proofs or certify
literature novelty.
