\section{Finite-pointer entropy and observer-code accuracy}
\label{sec:finite-pointer}

We now compose the sharp binary covariance theorem with the exact
finite-pointer channel.  The composition uses no new spectral estimate.  Its
input is the pairwise form bound; its additional ingredients are the weighted
variance identity and convexity.

Define the ideal classical pointer purity and its second R\'enyi entropy by
\begin{equation}
 \Pcl:=\sum_iw_i^2,
 \qquad H_2(w):=-\log\Pcl.
 \label{eq:classical-pointer-purity}
\end{equation}

\begin{theorem}[Finite-pointer energy--support--entropy bound]
\label{thm:finite-pointer-renyi}
Let $p_i$ be Dirichlet half-line momentum data supported in $[0,L]$, with
$w_i\geq0$ and $\sum_iw_i=1$.  Set
\begin{equation}
 \bar p=\sum_iw_ip_i,
 \qquad
 \Ebar=\frac12\sum_iw_i\norm{p_i-\bar p}^2.
 \label{eq:half-line-centered-energy}
\end{equation}
For the exact thermal channel in \cref{eq:dephasing-channel},
\begin{equation}
 \boxed{
 \Tr\rho_P^2\geq\Pcl+(1-\Pcl)
 \exp\!\left[-\frac{\mathcal C_\beta(L)\Ebar}{1-\Pcl}\right]}
 \label{eq:finite-pointer-purity-bound}
\end{equation}
when $\Pcl<1$, with the continuous value one when $\Pcl=1$.  Consequently,
\begin{align}
 S_2(\rho_P)
 &\leq-\log\left\{\Pcl+(1-\Pcl)
 \exp\!\left[-\frac{\mathcal C_\beta(L)\Ebar}{1-\Pcl}\right]\right\}
 \label{eq:finite-pointer-entropy-bound}\\
 &\leq\min\{H_2(w),\mathcal C_\beta(L)\Ebar\}.
 \label{eq:simple-entropy-bound}
\end{align}
The first inequality is saturated for equal binary weights and opposite
top-eigenfunction profiles.  Global sharpness for arbitrary pointer dimension
is not asserted.
\end{theorem}

\begin{proof}
Write $d_{ij}=p_i-p_j$ and $x_{ij}=2\Gamma_{ij}$.  By
\cref{eq:pairwise-dephasing,eq:general-cost-definition},
\begin{equation}
 x_{ij}=\frac12\ip{d_{ij}}{B_{\beta,L}d_{ij}}
 \leq\frac{\mathcal C_\beta(L)}4\norm{d_{ij}}^2.
 \label{eq:pairwise-exponent-bound}
\end{equation}
The weighted variance identity gives
\begin{equation}
 \sum_{i,j}w_iw_j\norm{p_i-p_j}^2=4\Ebar,
 \label{eq:pairwise-variance-identity}
\end{equation}
and hence
\begin{equation}
 \sum_{i,j}w_iw_jx_{ij}\leq\mathcal C_\beta(L)\Ebar.
 \label{eq:average-exponent-bound}
\end{equation}
The terms $i=j$ have $x_{ii}=0$ and total weight $\Pcl$.  Normalize the
remaining weights by $1-\Pcl$ and apply Jensen's inequality to $e^{-x}$ in
\cref{eq:finite-pointer-purity}; this proves
\cref{eq:finite-pointer-purity-bound}.  The first simple bound follows because
the right side is at least $\Pcl$.  For $z\geq0$, convexity also gives
\begin{equation}
 \Pcl+(1-\Pcl)e^{-z/(1-\Pcl)}\geq e^{-z}.
\end{equation}
Taking $z=\mathcal C_\beta(L)\Ebar$ proves the second bound in
\cref{eq:simple-entropy-bound}.

For two equal weights and $p_\pm=\pm p$, the off-diagonal exponents coincide.
If $p$ is the top eigenfunction of $B_{\beta,L}$, both the form bound and
Jensen's inequality are equalities.  For more pointer states, equality would
require all nonzero differences to lie in the simple top eigenspace and to
have equal exponent.  We do not claim that those conditions can be met.
\end{proof}

The same argument works in the full static-patch phase space because the
energy is a Hilbert norm and \cref{thm:main} bounds the thermal form of every
difference datum.

\begin{corollary}[Conformal de Sitter finite-pointer bound]
\label{cor:de-sitter-finite-pointer}
Let every conditional angular and canonical datum be supported in $[0,yR]$,
and define $\Ebar$ by centering in the full energy norm as in
\cref{eq:centered-field-energy}.  Then
\begin{align}
 \Tr\rho_P^2
 &\geq\Pcl+(1-\Pcl)
 \exp\!\left[-\frac{R C_{\rm opt}(y)\Ebar}{1-\Pcl}\right],
 \label{eq:de-sitter-pointer-purity}\\
 S_2(\rho_P)&\leq\min\{H_2(w),R C_{\rm opt}(y)\Ebar\}.
 \label{eq:de-sitter-pointer-entropy}
\end{align}
The coefficient $C_{\rm opt}(y)=2y\Lambda(y/2)$ is exact.  Binary saturation
uses the unique s-wave momentum optimizer of \cref{thm:main}.
\end{corollary}

\subsection{Insertion into the observer code}

The field record provides a purification of the physical pointer state:
\begin{equation}
 \ket\omega=\sum_i\sqrt{w_i}\ket{i}_{\rm Ob}\ket{e_i}_{\rm Ob'},
 \qquad \ip{e_j}{e_i}=G_{ij}.
 \label{eq:physical-record-state}
\end{equation}
The two reductions of this pure state have identical nonzero spectra, so
\begin{equation}
 e^{-S_2(\omega_{\rm Ob'})}
 =\Tr\omega_{\rm Ob'}^2
 =\Tr\rho_P^2.
 \label{eq:physical-purity-identification}
\end{equation}
This is the entropy parameter appearing in the simple random-code calculation
of Harlow, Usatyuk, and Zhao~\cite{HarlowUsatyukZhao2025}.

\begin{proposition}[Physical floor in the simple observer code]
\label{prop:harlow-floor}
Let $D$ be the dimension of their Haar-random orthogonal encoding matrix, and
let $\ket\phi$, $\ket\psi$ be orthogonal CRT-real matter states.  Replacing
the ideal record by \cref{eq:physical-record-state}, their exact second-moment
formula gives
\begin{equation}
 \mathbb E_O\abs{\bra\phi\widehat V^\dagger\widehat V\ket\psi}^2
 =\frac{D}{D+2}\Tr\rho_P^2.
 \label{eq:harlow-exact-purity}
\end{equation}
Consequently, for the de Sitter model,
\begin{equation}
 \mathbb E_O\abs{\bra\phi\widehat V^\dagger\widehat V\ket\psi}^2
 \geq\frac{D}{D+2}\left\{\Pcl+(1-\Pcl)
 e^{-R C_{\rm opt}(y)\Ebar/(1-\Pcl)}\right\}.
 \label{eq:harlow-physical-floor}
\end{equation}
\end{proposition}

\begin{proof}
Equation~(4.2) of \cite{HarlowUsatyukZhao2025} contains the observer terms
$e^{-S_2(\omega_{\rm Ob'})}$ and
$\Tr(\omega_{\rm Ob}\omega_{\rm Ob}^{T})$ together with two matter overlaps.
For the stated orthogonal CRT-real pair, both overlaps vanish, leaving
\cref{eq:harlow-exact-purity}.  Apply
\cref{eq:physical-purity-identification,eq:de-sitter-pointer-purity}.
\end{proof}

The proposition closes a specific gap: the entropy suppressing the simple
code fluctuation is now generated by a localized field record with an exact
energy--support bound.  Its scope is narrower than a gravitational observer
theorem.  The system called ${\rm Ob'}$ in
\cref{eq:physical-record-state} is the purified nonideal record, and
\cref{eq:harlow-physical-floor} is a Haar-ensemble mean-square statement for
one orthogonal pair.  It is not a deterministic lower bound for every fixed
encoding map.  Nor does it include an autonomous observer, actuator work, or
gravitational evolution.
