\section{Introduction}
\label{sec:introduction}

The observer rule of Harlow, Usatyuk, and Zhao replaces the inner product of
closed-universe states by a quantum-to-classical channel carried by an
observer record~\cite{HarlowUsatyukZhao2025}.  Its ideal motivation uses an
exact pointer-basis clone.  In their simple random-code model, departures from
an ideal inner product are controlled by the purity
$e^{-S_2(\omega_{\rm Ob'})}$ of the record placed in the corresponding
bipartite slot.  This makes the second R\'enyi entropy operational, but leaves
a complementary physical question: what happens when that ideal record is
replaced by a localized, nonideal field record of specified purity?

We answer that question in an exactly solvable model.  A finite pointer
controls compact classical sources for a free scalar field in a thermal
static patch.  The exact time-ordered evolution is a controlled Weyl
displacement, so tracing the field gives a Schur dephasing channel.  If
$w_i$ is the initial pointer distribution and $p_i$ are its conditional
post-switch momentum data, define
\begin{equation}
 \Pcl=\sum_iw_i^2,
 \qquad
 \bar p=\sum_iw_ip_i,
 \qquad
 \Ebar=\frac12\sum_iw_i\norm{p_i-\bar p}^2.
 \label{eq:intro-centered-data}
\end{equation}
Here $\Pcl$ is the purity of the ideal classical distribution and $\Ebar$ is
the field energy that distinguishes the conditional records; a common
displacement costs energy but carries no pointer information.

Suppose every conditional datum is supported in the half-line interval
$[0,L]$.  Our central finite-pointer result is
\begin{align}
 \Tr\rho_P^2
 &\geq \Pcl+(1-\Pcl)
 \exp\!\left[-\frac{\mathcal C_\beta(L)\Ebar}{1-\Pcl}\right],
 \label{eq:intro-purity-bound}\\
 S_2(\rho_P)
 &\leq\min\{H_2(w),\mathcal C_\beta(L)\Ebar\}.
 \label{eq:intro-renyi-bound}
\end{align}
The continuous value is understood when $\Pcl=1$.  The coefficient is the
exact localized norm of the thermal inverse Hamiltonian,
\begin{equation}
 \mathcal C_\beta(L)=2L\Lambda(\pi L/\beta),
 \qquad
 k_\tau(u,v)=\frac1\pi
 \log\frac{\sinh[\tau(u+v)]}{\sinh[\tau\abs{u-v}]},
 \label{eq:intro-kernel}
\end{equation}
where $\Lambda(\tau)$ is the simple top eigenvalue of the positive compact
operator with kernel $k_\tau$ on $L^2(0,1)$.  The purity bound follows by
combining the sharp pairwise form estimate with a weighted variance identity
and Jensen's inequality.  Equal binary weights and opposite top-eigenfunction
profiles saturate it.  We do not claim global sharpness for arbitrary pointer
dimension.

The operator theorem underlying \cref{eq:intro-kernel} is independent of de
Sitter geometry.  For Dirichlet half-line momentum data,
\begin{equation}
 \ip{p}{h_0^{-1}\coth(\beta h_0/2)p}
 \leq \frac12\mathcal C_\beta(L)\norm p^2,
 \label{eq:intro-half-line-bound}
\end{equation}
and the exact coefficient obeys the global bracket
\begin{equation}
 \max\left\{\frac{3L}{\pi},\frac{16L^2}{\beta\pi^2}\right\}
 \leq\mathcal C_\beta(L)
 \leq\frac{4\operatorname{arsinh}(1)}\pi L
      +\frac{16L^2}{\beta\pi^2}.
 \label{eq:intro-general-bracket}
\end{equation}
Uniform remainders identify the vacuum-controlled small-support law and the
thermal large-support law.

We then specialize to the conformally coupled massless scalar in a
four-dimensional de Sitter static patch.  The Bunch--Davies temperature has
$\beta=2\pi R$, and the optical partial-wave operators are
\begin{equation}
 A_\ell=-\partial_x^2+\frac{\ell(\ell+1)}
 {R^2\sinh^2(x/R)}.
\end{equation}
Resolvent order shows that no higher angular momentum has a larger thermal
momentum form.  A separate localized estimate controls the field-coordinate
sector.  Consequently, the half-line momentum optimizer is the unique
optimum over the full angular and canonical phase space, with
\begin{equation}
 \mathcal C_{2\pi R}(L)=R C_{\rm opt}(y),
 \qquad C_{\rm opt}(y)=2y\Lambda(y/2),
 \qquad y=L/R.
 \label{eq:intro-de-sitter-cost}
\end{equation}

We do not identify the field record with the exact pointer-basis clone used to
motivate the ideal observer rule.  We replace that ideal record with the
Stinespring record of a partial dephasing channel and ask how its physically
constrained purity modifies the same random-code second moment.  Purifying the
conditional field states gives
\begin{equation}
 \ket\omega=\sum_i\sqrt{w_i}\ket i_{\rm Ob}\ket{e_i}_{\rm Ob'},
 \qquad e^{-S_2(\omega_{\rm Ob'})}=\Tr\rho_P^2.
 \label{eq:intro-record-state}
\end{equation}
For an orthogonal CRT-real matter pair, substituting this state into the exact
second-moment formula of Harlow, Usatyuk, and Zhao yields
\begin{equation}
 \mathbb E_O
 \abs{\bra\phi\widehat V^\dagger\widehat V\ket\psi}^2
 =\frac{D}{D+2}\Tr\rho_P^2,
 \label{eq:intro-harlow-insertion}
\end{equation}
so \cref{eq:intro-purity-bound} becomes a field-energy lower floor on this
Haar-averaged code fluctuation.  This is an ensemble mean-square statement
for the specified pair, not a deterministic bound for every fixed code.

Finally, we give a limited gravitational application.  If every conditional
spherical $q_i=0$ branch obeys the final-slice local constraint budget
$Q_{b,i}\leq\delta$, then
\begin{equation}
 S_2(\rho_P)\leq
 \delta\frac{R^2}{G}\,
 \frac{C_{\rm opt}(y)\tanh y\,\sech^2y}{2}.
 \label{eq:intro-gravity-bound}
\end{equation}
The budget must hold branchwise: controlling only an averaged or centered
source does not control the conditional geometries.  This result composes a
fixed-background channel calculation with exact Einstein--scalar
final-slice constraints.  It neither recomputes the channel on the
branch-dependent geometries nor solves a coupled evolution.

Compact final support is essential throughout.  Without it, arbitrarily soft
thermal modes prevent the energy norm from controlling the dephasing form.
If a source is confined to areal radius $a<R$ for a static-time interval of
length $T$, finite propagation supplies the causal envelope
\begin{equation}
 L=R\operatorname{arctanh}(a/R)+T.
 \label{eq:intro-source-envelope}
\end{equation}
Smooth compact sources approach the final-support optimum when their
worldtube approaches the final-support ball.  We do not claim that every
fixed smaller source cylinder can approach the top mode.

The field resource in this paper is equally specific.  It is the
post-switch scalar Killing energy of the prescribed conditional sources.  It
does not include a clock, trap, battery, actuator, or autonomous observer
stress.  Exact gapless-detector channels and their energy accounting are
prior art~\cite{Landulfo2016,BarcellosLandulfo2021,BatistaEtAl2026}; positive
compact kernels and logarithmic-potential variational principles also have a
substantial literature.  The candidate contribution is the finite-pointer
purity theorem with the exact localized thermal coefficient, its all-sector
de Sitter reduction, and the controlled substitution of that physical record
in the observer-code second moment.  External detector/QFT, operator-theory,
and quantum-gravity review remain publication gates.

\Cref{sec:model} defines the controlled-source model and exact Schur channel.
\Cref{sec:bound} proves the localized thermal operator theorem and the full
de Sitter reduction.  \Cref{sec:finite-pointer} derives the R\'enyi bound and
observer-code record substitution.  \Cref{sec:construction} treats smooth source
realization, \cref{sec:comparison} separates the result from adjacent entropy
and detector bounds, and \cref{sec:discussion} states the claim boundary.
\Cref{sec:gravity} contains the branchwise final-slice application.
