\section{Static-patch model and exact pointer channel}
\label{sec:model}

\subsection{Optical partial waves}

In static coordinates, four-dimensional de Sitter space has
\begin{equation}
 \dd s^2=-N(r)\dd t^2+\frac{\dd r^2}{N(r)}+r^2\dd\Omega_2^2,
 \qquad N(r)=1-\frac{r^2}{R^2},
 \qquad 0\leq r<R .
 \label{eq:static-metric}
\end{equation}
Introduce the optical radius
\begin{equation}
 x=R\operatorname{arctanh}(r/R),
 \qquad r=R\tanh(x/R),
 \qquad \Omega(x)=\sech(x/R).
 \label{eq:optical-radius}
\end{equation}
Then $g_{\rm dS}=\Omega^2g_{\rm opt}$, where
\begin{equation}
 g_{\rm opt}=-\dd t^2+\dd x^2+R^2\sinh^2(x/R)\dd\Omega_2^2 .
\end{equation}
For the conformally coupled massless scalar, spherical-harmonic decomposition
and the standard radial rescaling give the positive operators
\begin{equation}
 A_\ell=-\frac{\dd^2}{\dd x^2}
 +\frac{\ell(\ell+1)}{R^2\sinh^2(x/R)},
 \qquad h_\ell=A_\ell^{1/2}
 \label{eq:partial-wave-hamiltonian}
\end{equation}
on $L^2(\R_+)$, with the regular Dirichlet condition at $x=0$.
The Bunch--Davies state restricted to the static patch is KMS at
\begin{equation}
 \beta=2\pi R
 \label{eq:beta}
\end{equation}
for the static Killing flow.  The half-line theorem in \cref{sec:bound} first
keeps $\beta$ arbitrary; \cref{thm:main} then imposes this de Sitter relation
when comparing all angular and canonical sectors.

Let $(q_{\ell m},p_{\ell m})$ denote real final Cauchy data for the rescaled
partial waves.  For smooth compact data, the thermal quadratic form and the
Killing energy are
\begin{align}
 \Gamma_{\ell m}
 &=\ip{q_{\ell m}}{h_\ell\coth(\beta h_\ell/2)q_{\ell m}}
   +\ip{p_{\ell m}}{h_\ell^{-1}\coth(\beta h_\ell/2)p_{\ell m}},
 \label{eq:gamma-sector}\\
 E_{\ell m}
 &=\frac12\left(\norm{h_\ell q_{\ell m}}^2
                    +\norm{p_{\ell m}}^2\right).
 \label{eq:energy-sector}
\end{align}
We write $\Gamma=\sum_{\ell m}\Gamma_{\ell m}$ and
$\EK=\sum_{\ell m}E_{\ell m}$.  Smooth angular data make these sums finite in
the needed form domains.

For $\ell=0$, the unitary sine transform diagonalizes $h_0$.  With hats
denoting that transform,
\begin{align}
 \Gamma_{00}
 &=\int_0^\infty\coth(\beta k/2)
 \left(k\abs{\widehat q(k)}^2+\frac{\abs{\widehat p(k)}^2}{k}\right)\dd k,
 \label{eq:gamma-sine}\\
 E_{00}
 &=\frac12\int_0^\infty
 \left(k^2\abs{\widehat q(k)}^2+\abs{\widehat p(k)}^2\right)\dd k.
 \label{eq:energy-sine}
\end{align}

\subsection{Local interaction and exact channel}

Let $P$ be a finite pointer with orthonormal basis
$\{\ket{i}\}_{i=1}^d$.  Assign a real compact source $J_i$ to each pointer
state and couple it to the physical scalar by
\begin{equation}
 S_{\rm int}=-\int_{\mathcal M}\sqrt{-g}\,\phi(X)
 \sum_{i=1}^dJ_i(X)\ket{i}\!\bra{i}\,\dd^4X,
 \qquad J_i\in C_c^\infty(\mathcal M;\R).
 \label{eq:interaction}
\end{equation}
The switching functions and spatial smearings are included in the $J_i$.
The pointer projectors commute, and free-field commutators are c-numbers, so
the Magnus series terminates in every controlled sector.  Up to
source-dependent phases,
\begin{equation}
 U=\sum_{i=1}^d e^{i\Xi[J_i]}\ket{i}\!\bra{i}
 \otimes e^{-i\phi(J_i)}.
 \label{eq:exact-unitary}
\end{equation}
This standard exact evolution is reviewed in
\cite{Landulfo2016,BarcellosLandulfo2021,BatistaEtAl2026}.
The interaction is compactly supported, and its densities commute at
spacelike separation because the diagonal pointer operators commute and the
scalar field is microcausal.  Nevertheless, the same finite pointer appears
across each spatial smearing.  Generic spatially extended nonrelativistic
detectors can violate causal factorization~\cite{deRamonEtAl2021}; fully
relativistic localized probe fields provide a systematic replacement
framework~\cite{PercheEtAl2024}.  We treat \cref{eq:interaction} as a
prescribed gapless-detector idealization, not as an autonomous pointer field
or material actuator.

Let $E$ denote the causal propagator and $K$ the positive-frequency map for the
static one-particle structure.  Purify the Bunch--Davies KMS state and let
$\ket{e_i}$ be the conditional purified field state.  Their Gram matrix
$G_{ij}=\ip{e_j}{e_i}$ is positive semidefinite with unit diagonal.  Tracing
the field gives the exact Schur dephasing channel
\begin{equation}
 \cD_G(\ket{i}\!\bra{j})=G_{ij}\ket{i}\!\bra{j},
 \qquad
 \abs{G_{ij}}=e^{-\Gamma_{ij}}.
 \label{eq:dephasing-channel}
\end{equation}
Let $(q_i,p_i)$ be the complete conditional final data, and write
$q_{ij}=q_i-q_j$, $p_{ij}=p_i-p_j$.  Gaussianity gives
\begin{equation}
 \Gamma_{ij}=\frac14\sum_{\ell m}\left[
 \ip{q_{ij,\ell m}}{h_\ell\coth(\beta h_\ell/2)q_{ij,\ell m}}
 +\ip{p_{ij,\ell m}}{h_\ell^{-1}\coth(\beta h_\ell/2)p_{ij,\ell m}}
 \right].
 \label{eq:pairwise-dephasing}
\end{equation}
For an initial pointer state
$\ket\psi=\sum_i\sqrt{w_i}\ket i$, the reduced pointer state is
\begin{equation}
 \rho_{P,ij}=\sqrt{w_iw_j}\,G_{ij},
 \qquad
 \Tr\rho_P^2=\sum_{i,j}w_iw_j e^{-2\Gamma_{ij}}.
 \label{eq:finite-pointer-purity}
\end{equation}

The binary model is recovered by $d=2$, $J_\pm=\pm J$.  Its conditional data
are opposite, so \cref{eq:pairwise-dephasing} gives
$\Gamma_{+-}=\Gamma$ with
\begin{equation}
 \Gamma=2\ip{KEJ}{\coth(\beta h/2)KEJ}.
\end{equation}
The phase of $G_{+-}=\kappa$ is a correctable pointer rotation.  If $\cD_0$
denotes complete binary dephasing, then
\begin{equation}
 \eobs:=\frac12\norm{\cD_\kappa-\cD_0}_\diamond
 =\frac12\abs\kappa=\frac12e^{-\Gamma}.
 \label{eq:diamond-error}
\end{equation}
The equality follows from
$\cD_\kappa-\cD_0=(\kappa/2)(\mathrm{id}-\operatorname{Ad}_{Z_P})$ after the
correctable phase is removed, and it is saturated by a pointer superposition.
The exact quantum-to-classical channel used in the observer rule of Harlow,
Usatyuk, and Zhao~\cite{HarlowUsatyukZhao2025} is the limit in which all
off-diagonal $G_{ij}$ vanish.  Rather than identify the binary $\eobs$ with a
gravitational encoding error, \cref{sec:finite-pointer} below uses the
physical purity in \cref{eq:finite-pointer-purity}, which is the quantity that
enters their second-R\'enyi code calculation.

After switching, branch $i$ has excess scalar Killing energy $E_i$ and
renormalized-stress excess
\begin{equation}
 E_i=\frac12\sum_{\ell m}
 \left(\norm{h_\ell q_{i,\ell m}}^2+\norm{p_{i,\ell m}}^2\right),
 \qquad
 \Delta\langle T_{ab}^{\rm ren}\rangle_i=T_{ab}[\phi_{J_i}].
 \label{eq:energy-stress}
\end{equation}
Let $\bar q=\sum_iw_iq_i$, $\bar p=\sum_iw_ip_i$.  The energy that controls
pairwise separation is
\begin{equation}
 \Ebar:=\sum_iw_i E[q_i-\bar q,p_i-\bar p]
 =\sum_iw_iE_i-E[\bar q,\bar p].
 \label{eq:centered-field-energy}
\end{equation}
For the symmetric binary model, $\bar q=\bar p=0$ and
$E_+=E_-=\Ebar=\EK$.
The state-independent Hadamard subtraction cancels in the difference, and the
background one-point function vanishes.  These quantities price the
post-switch scalar fields generated by the same $J_i$ as the channel.  They do
not include the stress of a clock, trap, or material device that realizes the
prescribed coupling.

\subsection{Causal support}

Suppose every $J_i$ is supported inside the centered areal ball $r\leq a<R$
during a common static-time interval of length $T$.  Its optical radius is
\begin{equation}
 \ell=R\operatorname{arctanh}(a/R).
 \label{eq:source-optical-radius}
\end{equation}
Finite propagation in the optical metric implies that the final Cauchy data of
every partial wave are supported in
\begin{equation}
 0\leq x\leq L,
 \qquad L=\ell+T.
 \label{eq:causal-support}
\end{equation}
No spectral cutoff or hard spatial wall is introduced.
