\documentclass[11pt]{article}

\usepackage[T1]{fontenc}
\usepackage{newtxtext,newtxmath}
\usepackage[margin=1.08in]{geometry}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{enumitem}
\usepackage{microtype}
\usepackage[hidelinks]{hyperref}
\usepackage[nameinlink,noabbrev]{cleveref}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

\newcommand{\cH}{\mathcal H}
\newcommand{\cC}{\mathcal C}
\newcommand{\cK}{\mathcal K}
\newcommand{\1}{\mathbf 1}
\newcommand{\Tr}{\operatorname{Tr}}
\newcommand{\rank}{\operatorname{rank}}
\newcommand{\SO}{\mathrm{SO}(3)}
\newcommand{\supp}{\operatorname{supp}}
\newcommand{\Rref}{R_{\mathrm{ref}}}
\newcommand{\positive}[1]{\left[#1\right]_{+}}

\title{Methods Note: Off-Code Leakage from Spacelike Replication\\
of a Finite $\mathrm{SO}(3)$ Collective Mode}
\author{Brian Fioca}
\date{Internal methods note, June 2026}

\begin{document}
\maketitle

\begin{center}
\small
\textbf{NOVELTY STOP.}
The bounded-operator results are correct, but the main inequality is a direct
specialization of Janssens' CP-map covariance Cauchy--Schwarz lemma. This note
is retained as an explicit $\SO$ application and internal reference, not as a
standalone paper claim.
\end{center}

\begin{abstract}
We record a finite-code specialization of established joint-measurement
added-noise bounds. Three bounded observables localized in
pairwise spacelike regions commute microscopically but compress to the three
noncommuting generators of a single finite $\SO$ collective mode. If $P$ is
the code projector, $Q=\1-P$, $P A_aP=\alpha J_a$, and
$\Lambda\geq\max_a\lVert Q A_aP\rVert$ is a certified off-code amplitude
cap, then every code state $\rho$ obeys
\[
  \sum_a \Tr(\rho P A_aQ A_aP)
  \geq \frac{\alpha^4\Tr(\rho J^2)}{4\Lambda^2}.
\]
The pairwise state-weighted inequality follows directly from the covariance
Cauchy--Schwarz lemma for the UCP compression $T(X)=W^*XW$; cyclic summation
gives the displayed $\SO$ form. We also record an explicit robust version and
compose the exact inequality with a global Haar-prior orientation-risk bound.
Disjoint blocks in the symmetric sector of a ferromagnetic spin chain make
the underlying pairwise norm coefficient asymptotically sharp and realize
the three-cell scaling within a factor tending to eight. The result is an
absolute fixed-calibration statement: because
$\lVert\alpha J_a\rVert\leq\lVert A_a\rVert\leq M$, the dimensionless gain
$g=|\alpha|/M$ satisfies $g\leq 1/J$ on a spin-$J$ code, so relative leakage
may vanish at large spin. No transition rate, lifetime, uniformly bounded
support, unbounded-current theorem, or gravitational conclusion is inferred.
\end{abstract}

\section{Introduction}
\label{sec:introduction}

Compression can turn commuting microscopic observables into noncommuting
logical ones. For a unital completely positive map, the associated covariance
form and its Cauchy--Schwarz inequality quantify exactly this added noise
\cite{janssens2017,polterovich2014}. If $P$ is a projector and $Q=\1-P$,
inserting $P+Q$ between two operators gives the corresponding compression
identity directly. Related mechanisms are standard in Toeplitz quantization
and quantum error correction \cite{bordemann1994,beny2009,faist2020}. The
purpose of this note is to spell out the following $\SO$ application:

\begin{quote}
What off-code cost is forced when three mutually commuting cells are each
required to reproduce a different component of one rigid finite $\SO$
collective mode?
\end{quote}

The distinction between one region and several commuting regions is
essential. Locality does not require different non-Abelian current components
inside one region to commute, and the results below are not a no-go theorem
for local non-Abelian symmetry. They concern replication of distinct
components across distinct commuting cells after projection to one finite
collective code.

The main result is state weighted. For a code state $\rho$, define
\[
  p_a(\rho)=\Tr(\rho P A_aQ A_aP)
           =\lVert Q A_aP\rho^{1/2}\rVert_2^2.
\]
These quantities are squared off-code amplitudes. They are not transition
probabilities unless the $A_a$ are embedded into a normalized operation or a
dynamics. By summing the three cyclic $\mathfrak{so}(3)$ commutators, we bound
$\sum_a p_a$ below by the mean Casimir. A separate global estimation lemma
then converts achieved full-frame orientation risk into a necessary Casimir
and hence into off-code weight.

Three qualifications are part of the record. First, the gain $\alpha$ is
physical calibration data and cannot be rescaled away. Second, the explicit
ferromagnetic blocks grow with system size and the parent Hamiltonian is not
uniformly gapped. Third, both the compression identity and the state-weighted
inequality are standard CP-map/joint-measurement consequences. The exact
prior-art reduction is explicit in \cref{sec:related}; no standalone novelty
is claimed.

\section{Finite Code and Commuting Cells}
\label{sec:setup}

Let $\cH$ be a Hilbert space, let $P$ be a finite-rank orthogonal projector,
and put
\[
  \cC=P\cH,\qquad Q=\1-P.
\]
The code $\cC$ carries a finite-dimensional integer-spin representation of
$\SO$, possibly with rotation-trivial multiplicities. Its self-adjoint
generators satisfy
\[
  [J_a,J_b]=i\epsilon_{abc}J_c,
  \qquad J^2=J_x^2+J_y^2+J_z^2.
\]
Write $J_{\max}$ for the largest spin present in the code.

For the local-net formulation, choose three pairwise spacelike regions
$O_a$ and bounded self-adjoint observables $A_a\in\mathcal A(O_a)$. Only the
consequence
\begin{equation}
  [A_a,A_b]=0\quad(a\neq b)
  \label{eq:cell-commutation}
\end{equation}
is used. Thus the same proof applies to any three pairwise commuting bounded
star-algebras, including disjoint cells of a finite lattice.

We use two amplitude scales:
\begin{equation}
  \lambda_*:=\max_a\lVert Q A_aP\rVert,
  \qquad \lambda_*\leq\Lambda,
  \qquad 0<\Lambda\leq M,
  \qquad \lVert A_a\rVert\leq M.
  \label{eq:amplitude-scales}
\end{equation}
Here $\Lambda$ is a caller-supplied certified cap. The coarse choice
$\Lambda=M$ is always available, but a smaller cap must be established
independently.

\section{Compression and State-Weighted Leakage}
\label{sec:main}

We begin with the standard compression identity to fix signs and constants.

\begin{lemma}[Compression identity]
\label{lem:compression}
For bounded operators $A,B$,
\begin{equation}
  [PAP,PBP]
  =P[A,B]P+PBQAP-PAQBP.
  \label{eq:compression-identity}
\end{equation}
If $A$ and $B$ are self-adjoint, then
\begin{equation}
  \lVert[PAP,PBP]\rVert
  \leq \lVert P[A,B]P\rVert
       +2\lVert QAP\rVert\lVert QBP\rVert.
  \label{eq:pairwise-norm}
\end{equation}
\end{lemma}

\begin{proof}
Expand $PABP$ and $PBAP$ after inserting $\1=P+Q$ and subtract. For
self-adjoint operators,
$\lVert PAQ\rVert=\lVert QAP\rVert$ and similarly for $B$, so the triangle
inequality gives \cref{eq:pairwise-norm}.
\end{proof}

The coefficient two in \cref{eq:pairwise-norm} cannot be reduced uniformly;
\cref{sec:model} gives a disjoint-block family whose ratio tends to one.

\begin{theorem}[State-weighted spacelike-replication leakage]
\label{thm:three-cell}
Let $A_x,A_y,A_z$ be bounded self-adjoint operators satisfying
\cref{eq:cell-commutation,eq:amplitude-scales}, and suppose that for a real
gain $\alpha$,
\begin{equation}
  P A_aP=\alpha J_a\qquad(a=x,y,z).
  \label{eq:exact-compression}
\end{equation}
Then every density operator $\rho=P\rho P$ obeys
\begin{equation}
  \alpha^4\Tr(\rho J^2)
  \leq 4\lambda_*^2\sum_a p_a(\rho)
  \leq 4\Lambda^2\sum_a p_a(\rho),
  \label{eq:direct-product-form}
\end{equation}
and therefore
\begin{equation}
  \sum_a p_a(\rho)
  \geq \frac{\alpha^4\Tr(\rho J^2)}{4\Lambda^2}
  \geq \frac{\alpha^4\Tr(\rho J^2)}{4M^2}.
  \label{eq:main-bound}
\end{equation}
\end{theorem}

\begin{proof}
For a cyclic triple $(a,b,c)$, \cref{lem:compression}, microscopic
commutation, and \cref{eq:exact-compression} give
\begin{equation}
  i\alpha^2J_c=P A_bQ A_aP-P A_aQ A_bP.
  \label{eq:cyclic-factorization}
\end{equation}
Right multiplication by $\rho^{1/2}$ followed by the Hilbert--Schmidt
triangle inequality yields
\begin{align*}
  |\alpha|^2\sqrt{\Tr(\rho J_c^2)}
  &\leq \lVert P A_bQ\rVert
          \lVert Q A_aP\rho^{1/2}\rVert_2
       + \lVert P A_aQ\rVert
          \lVert Q A_bP\rho^{1/2}\rVert_2 \\
  &\leq \lambda_*\bigl(\sqrt{p_a(\rho)}+\sqrt{p_b(\rho)}\bigr).
\end{align*}
Squaring and using $(u+v)^2\leq2(u^2+v^2)$ gives
\[
  \alpha^4\Tr(\rho J_c^2)
  \leq2\lambda_*^2\bigl(p_a(\rho)+p_b(\rho)\bigr).
\]
Summing the three cyclic inequalities counts each $p_a$ twice and proves
\cref{eq:direct-product-form,eq:main-bound}.
\end{proof}

\begin{corollary}[Consistency of a certified cap]
\label{cor:cap-consistency}
Under the hypotheses of \cref{thm:three-cell},
\begin{equation}
  \Lambda^4\geq\frac{\alpha^4\Tr(\rho J^2)}{12}.
  \label{eq:cap-consistency}
\end{equation}
\end{corollary}

\begin{proof}
Each $p_a\leq\lVert Q A_aP\rVert^2\leq\Lambda^2$, so
$\sum_a p_a\leq3\Lambda^2$. Combine this upper bound with
\cref{eq:main-bound}.
\end{proof}

\begin{remark}[Calibration and the large-spin escape]
\label{rem:calibration}
On a spin-$J$ code, exact compression implies
\[
  |\alpha|J=\lVert\alpha J_a\rVert
  =\lVert P A_aP\rVert\leq\lVert A_a\rVert\leq M.
\]
Consequently $g=|\alpha|/M\leq1/J$. The normalized form of
\cref{eq:main-bound},
\[
  \frac{\sum_a p_a}{M^2}
  \geq \frac{g^4}{4}\Tr(\rho J^2),
\]
can therefore weaken at large spin. The theorem is an absolute
fixed-calibration obstruction, not a statement that better references must
decohere faster.
\end{remark}

For completeness, the uniform approximate statement is recorded here and
proved in \cref{app:robust}.

\begin{proposition}[Robust three-cell bound]
\label{prop:robust}
Let $A_x,A_y,A_z$ be bounded self-adjoint operators, let
$\alpha\in\mathbb R$, $\delta,\epsilon\geq0$, and $\Lambda>0$. Suppose that
\[
  \lVert P[A_a,A_b]P\rVert\leq\delta\quad(a\neq b),
  \qquad
  \lVert P A_aP-\alpha J_a\rVert\leq\epsilon,
  \qquad
  \lVert Q A_aP\rVert\leq\Lambda\quad(a=x,y,z).
\]
Define
\begin{equation}
  d=\delta+4|\alpha|J_{\max}\epsilon+2\epsilon^2.
  \label{eq:defect-d}
\end{equation}
Then, for every $t>0$ and every code state $\rho=P\rho P$,
\begin{equation}
  \sum_a p_a(\rho)
  \geq
  \frac{
    \positive{\alpha^4\Tr(\rho J^2)-3(1+t^{-1})d^2}
  }{4(1+t)\Lambda^2}.
  \label{eq:robust-bound}
\end{equation}
\end{proposition}

\section{Operational Orientation Risk}
\label{sec:risk}

Let an unknown $g\in\SO$ have normalized Haar prior. An arbitrary POVM
returns an estimate $\hat g$, and the full-frame chordal cost is
\begin{equation}
  c(g,\hat g)
  =\sin^2\!\left(\frac{\theta(\hat g^{-1}g)}{2}\right)
  =\frac{3-\chi_1(\hat g^{-1}g)}{4},
  \label{eq:chordal-cost}
\end{equation}
where $\chi_1$ is the spin-one character. The Bayes risk optimized over all
measurements is denoted by $\Rref$. More explicitly, write
$R(\rho,\mathsf M)$ for the Bayes risk of a fixed POVM $\mathsf M$ and
$\Rref(\rho)=\inf_{\mathsf M}R(\rho,\mathsf M)$. The orbit experiment is
$\rho_g=U(g)\rho U(g)^*$: all unknown-$g$ dependence is carried by this orbit,
the POVM is fixed independently of $g$, and any additional side system is
rotation trivial. An extra $g$-dependent reference would be an additional
resource and is not included in the theorem.
This is the standard covariant full-frame estimation task
\cite{holevo1982,peres2001,bagan2001}.

\begin{theorem}[Global mean-Casimir risk floor]
\label{thm:risk-floor}
For every integer-spin $\SO$ reference state, with arbitrary
rotation-trivial multiplicities, and every orientation POVM $\mathsf M$,
\begin{equation}
  R(\rho,\mathsf M)\geq\frac{1}{16\Tr(\rho J^2)+8}.
  \label{eq:risk-floor}
\end{equation}
Consequently the same bound holds for $\Rref(\rho)$.
When $\Tr(\rho J^2)=\infty$, the right side is understood as zero.
\end{theorem}

The proof, including covariantization, arbitrary multiplicities and mixed
states, the Clebsch--Gordan block contraction, and the discrete Hardy step, is
given in \cref{app:risk}. The statement is global: it assumes neither local
unbiasedness nor a Cramer--Rao regime. It is a necessary resource law and does
not assert that large Casimir is sufficient for a good frame.

\begin{corollary}[Risk-conditioned off-code leakage]
\label{cor:risk-leakage}
Under the hypotheses of \cref{thm:three-cell}, if a protocol actually attains
$R(\rho,\mathsf M)\leq r$ with $0<r<1/8$, then
\begin{align}
  \Tr(\rho J^2)
  &\geq\frac{r^{-1}-8}{16},
  \label{eq:risk-casimir}\\
  \sum_a p_a(\rho)
  &\geq\frac{\alpha^4(r^{-1}-8)}{64\Lambda^2},
  \label{eq:risk-weight}\\
  \Lambda^4
  &\geq\frac{\alpha^4(r^{-1}-8)}{192}.
  \label{eq:risk-amplitude}
\end{align}
\end{corollary}

\begin{proof}
Invert \cref{eq:risk-floor}, then substitute the result into
\cref{eq:main-bound,eq:cap-consistency}.
\end{proof}

This implication is necessary, not sufficient. A declared finite spin support
must separately satisfy the sharper cutoff floor
\[
  \Rref\geq\sin^2\!\left(\frac{\pi}{2J_{\max}+3}\right),
\]
and compressed-action normalization requires
$|\alpha|J_{\max}\leq M$ in the exact case. Parameters violating either
condition describe no realization.

\section{Disjoint Ferromagnetic Realizations}
\label{sec:model}

Consider $N$ spin-one-half sites, with $N$ even, let $P_N$ project onto the
totally symmetric spin-$j=N/2$ subspace, and set $Q_N=\1-P_N$. This code is the
exact ground space of the nearest-neighbor frustration-free ferromagnet
\begin{equation}
  H_F=\sum_{i=1}^{N-1}\left(\frac14-S_i\cdot S_{i+1}\right).
  \label{eq:ferromagnet}
\end{equation}
Each summand projects onto the two-site singlet. A zero-energy vector is
therefore invariant under every adjacent transposition, whose common fixed
space is $\operatorname{Sym}^N(\mathbb C^2)$.

\begin{proposition}[Pairwise asymptotic sharpness]
\label{prop:pairwise-model}
Let $X,Y$ be disjoint blocks of sizes $r,s$ and define
\[
  A_x=\frac Nr\sum_{i\in X}S_i^x,
  \qquad
  B_y=\frac Ns\sum_{k\in Y}S_k^y.
\]
Then $[A_x,B_y]=0$ and
\[
  P_NA_xP_N=J_x,
  \qquad P_NB_yP_N=J_y.
\]
For complementary half-blocks,
\[
  \lVert Q_N A_xP_N\rVert
  =\lVert Q_N B_yP_N\rVert
  =\frac{j}{\sqrt{N-1}},
\]
and hence
\begin{equation}
  \frac{2\lVert Q_N A_xP_N\rVert\lVert Q_N B_yP_N\rVert}
       {\lVert J_z\rVert}
  =\frac{N}{N-1}\longrightarrow1.
  \label{eq:pairwise-sharp}
\end{equation}
Thus the coefficient two in \cref{eq:pairwise-norm} is optimal.
\end{proposition}

\begin{proposition}[Three-cell scaling]
\label{prop:three-cell-model}
Let $N$ be a multiple of six, divide the sites into three disjoint blocks
$X_a$ of size $N/3$, and put
\[
  A_a=3\sum_{i\in X_a}S_i^a,
  \qquad a=x,y,z.
\]
The three observables commute pairwise and compress exactly to $J_a$. Moreover,
\begin{align}
  \lambda_*^2
  &=\frac{N^2}{2(N-1)},
  \label{eq:model-lambda}\\
  p_a(\rho)
  &=\frac{2}{N-1}\left[j^2-\Tr(\rho J_a^2)\right],
  \label{eq:model-pa}\\
  \sum_a p_a(\rho)&=N.
  \label{eq:model-sum}
\end{align}
With the exact cap $\Lambda=\lambda_*$, the ratio of the actual weight to the
lower bound in \cref{eq:main-bound} is
\begin{equation}
  \frac{8N^2}{(N+2)(N-1)}\longrightarrow8.
  \label{eq:model-ratio}
\end{equation}
\end{proposition}

\begin{proof}[Proof of \cref{prop:pairwise-model,prop:three-cell-model}]
Permutation symmetry gives $P_NS_i^aP_N=J_aP_N/N$. For distinct sites
$i\neq k$, all ordered pairs are equivalent and
\[
  N(N-1)P_NS_i^aS_k^bP_N
  =P_N\left(J_aJ_b-\sum_{\ell}S_\ell^aS_\ell^b\right)P_N.
\]
Using
$S_\ell^aS_\ell^b=\delta_{ab}\1/4+i\epsilon_{abc}S_\ell^c/2$
and $[J_a,J_b]=i\epsilon_{abc}J_c$ yields
\begin{equation}
  P_NS_i^aS_k^bP_N
  =\frac{\{J_a,J_b\}/2-\delta_{ab}N\1/4}{N(N-1)}.
  \label{eq:pair-compression}
\end{equation}
Summing \cref{eq:pair-compression} over the declared blocks gives
\begin{align*}
  (Q_N A_xP_N)^*(Q_N A_xP_N)
  &=\frac{N-r}{r(N-1)}(j^2-J_x^2),\\
  P_NB_yQ_N A_xP_N
  &=\frac{\{J_x,J_y\}}{2(N-1)}+\frac{iJ_z}{2},
\end{align*}
and the analogous formulas with $x,y,z$ permuted. Their operator norms prove
\cref{prop:pairwise-model}. Setting $r=N/3$, taking expectations, and summing
over axes gives \cref{eq:model-lambda,eq:model-pa,eq:model-sum}; comparison
with \cref{eq:main-bound} gives \cref{eq:model-ratio}.
\end{proof}

The model establishes disjoint-support commutativity, not uniform locality.
The block diameters grow with $N$, and the ferromagnetic gap closes with system
size. No claim is made about local preparation, finite-time stability, a
continuum net, or a lifetime.

\section{Prior-Art Reduction and QEC Dictionary}
\label{sec:related}

The state-weighted inequality is already contained in the general CP-map
framework of Janssens \cite{janssens2017}. Let
$W:\mathcal C\to\mathcal H$ be the encoding isometry, so that
$W^*W=\1_{\mathcal C}$ and $WW^*=P$, and define
\[
  T(X)=W^*XW,
  \qquad
  (X,Y)_T=T(X^*Y)-T(X)^*T(Y).
\]
Write $\rho_{\mathcal C}=W^*\rho W$ and
$\lambda_a=\lVert Q A_aP\rVert=\lVert Q A_aW\rVert$. In the formulas below
we suppress the state subscript, equivalently identifying $\mathcal C$ with
its image $P\mathcal H$.
For a self-adjoint $A_a$,
\begin{equation}
  D_a:=(A_a,A_a)_T=W^*A_aQA_aW,
  \qquad
  \lVert D_a\rVert=\lambda_a^2,
  \qquad
  \Tr(\rho D_a)=p_a(\rho).
  \label{eq:noise-operator}
\end{equation}
Janssens' Lemma~1 proves
\[
  (X,Y)_T(Y,X)_T\leq\lVert(Y,Y)_T\rVert(X,X)_T.
\]
When $[A_a,A_b]=0$,
\[
  [T(A_a),T(A_b)]=(A_b,A_a)_T-(A_a,A_b)_T.
\]
The CP Cauchy--Schwarz inequality and the Hilbert--Schmidt triangle inequality
therefore give
\begin{equation}
  \lVert[T(A_a),T(A_b)]\rho^{1/2}\rVert_2
  \leq\lambda_b\sqrt{p_a(\rho)}
       +\lambda_a\sqrt{p_b(\rho)}.
  \label{eq:janssens-pair}
\end{equation}
After $T(A_a)=\alpha J_a$, this is exactly the pair inequality used in the
proof of \cref{thm:three-cell}. Squaring and summing its three cyclic
$\mathfrak{so}(3)$ instances gives the first inequality in
\cref{eq:direct-product-form}; $\lambda_*\leq\Lambda$ gives the certified-cap
version. Janssens' Theorem~3 states the corresponding pairwise
joint-measurement noise bound, and the construction following its proof
establishes sharpness. The proposed central theorem is therefore an application
of established added-noise theory, not a new obstruction. Polterovich uses the
same noise operator and inequality in the unsharpness principle
\cite{polterovich2014}. The joint spectral theorem directly supplies a common
PVM for the three commuting pointers used here. Related Naimark-dilation
characterizations cover pairs \cite{beneduci2014} and finite sets of
finite-outcome POVMs \cite{mitra2020}.

There is also an exact local-covariance dictionary:
\[
  A_aW-W\alpha J_a=Q A_aW.
\]
For a purification $|\psi_\rho\rangle$, define
\[
  |\Psi_a(t)\rangle
  =\bigl(e^{-itA_a}W e^{it\alpha J_a}\otimes\1\bigr)
   |\psi_\rho\rangle.
\]
Its tangent at zero is
$-i(Q A_aW\otimes\1)|\psi_\rho\rangle$ and is orthogonal to the encoded
state, so the pure-path QFI is
$F_a^{\mathrm{rel}}(\rho)=4p_a(\rho)$. This is not the QFI of the ordinary
orbit generated only by $A_a$, which also contains the logical variance.
Thus the exact-amplitude first inequality in \cref{eq:direct-product-form} is
equivalently
\begin{equation}
  \alpha^4\Tr(\rho J^2)
  \leq\lambda_*^2\sum_aF_a^{\mathrm{rel}}(\rho).
  \label{eq:qfi-form}
\end{equation}
This is a relative-covariance-QFI restatement of the same added-variance
operator. Nearby continuous-symmetry QEC and approximate-cleaning results
remain useful context
\cite{eastin2009,beny2009,flammia2017,faist2020,zhou2021,liuzhou2021}, but the
decisive reduction occurs already at the UCP/joint-measurement level.

\section{Discussion}
\label{sec:discussion}

The established added-noise theorem identifies a precise truncation cost in
this setting. If three commuting cells
are required to act inside one code as the three components of a rigid
non-Abelian collective mode, the omitted sector must mediate the logical
brackets. The state-weighted form shows that this mediation cannot be hidden
entirely in worst-case directions irrelevant to the chosen code state.

Several escape routes are physical rather than pathological:
\begin{enumerate}[leftmargin=2em]
\item place all noncommuting components in one region;
\item allow the local factors to leave the collective code;
\item reduce the gain as the spin grows;
\item use distributed modes not described by one rigid finite code; or
\item replace bounded observables by a separately controlled theorem for
unbounded affiliated operators.
\end{enumerate}

The explicit $\SO$ specialization and risk composition are retained as
internal tools; they do not support a standalone novelty claim. The result
also does not turn off-code weight into time. A lifetime statement would
require a Hamiltonian, a normalized interaction, and control of return as well
as departure from the code. Nor does the finite model establish a uniformly
localized apparatus. Those additions belong to the broader observer program,
not to the claim proved here.

\appendix

\section{Proof of the Robust Bound}
\label{app:robust}

Write $E_a=P A_aP-\alpha J_a$, so $\lVert E_a\rVert\leq\epsilon$. For a
cyclic triple $(a,b,c)$,
\begin{align*}
  \bigl\lVert[P A_aP,P A_bP]-i\alpha^2\epsilon_{abc}J_c\bigr\rVert
  &\leq
  \lVert[\alpha J_a,E_b]\rVert
  +\lVert[E_a,\alpha J_b]\rVert
  +\lVert[E_a,E_b]\rVert\\
  &\leq4|\alpha|J_{\max}\epsilon+2\epsilon^2.
\end{align*}
Combining this estimate with \cref{eq:compression-identity} and the locality
defect gives
\begin{equation}
  |\alpha|^2\sqrt{\Tr(\rho J_c^2)}
  \leq\Lambda(\sqrt{p_a}+\sqrt{p_b})+d.
  \label{eq:robust-pair}
\end{equation}
For $t>0$, $(x+y)^2\leq(1+t)x^2+(1+t^{-1})y^2$. Applying this to
\cref{eq:robust-pair}, then using
$(\sqrt{p_a}+\sqrt{p_b})^2\leq2(p_a+p_b)$ and summing cyclically, gives
\[
  \alpha^4\Tr(\rho J^2)
  \leq4(1+t)\Lambda^2\sum_a p_a
     +3(1+t^{-1})d^2.
\]
Rearrangement proves \cref{eq:robust-bound}.

\section{Global $\mathrm{SO}(3)$ Risk Proof}
\label{app:risk}

We prove \cref{thm:risk-floor} to make the operational corollary independent
of an internal repository result. Let
\[
  \cH=\bigoplus_{j\geq0}V_j\otimes\mathcal M_j,
  \qquad
  U(g)=\bigoplus_{j\geq0}U_j(g)\otimes\1_{\mathcal M_j},
\]
where the multiplicity spaces may be arbitrary and rotations act trivially on
them. Put $P_j$ for the spin-$j$ projector and
\[
  v_j=\sqrt{\Tr(P_j\rho)},
  \qquad \sum_jv_j^2=1,
  \qquad \overline C=\Tr(\rho J^2)=\sum_jj(j+1)v_j^2.
\]
If $\overline C=\infty$, \cref{eq:risk-floor} is trivial, so assume it is
finite.

\subsection*{Covariant reduction and fusion contraction}

Haar averaging turns any orientation POVM $\mathsf M$ into
\[
  \widetilde{\mathsf M}(B)
  =\int_{\SO}U(h)^*\mathsf M(hB)U(h)\,dh.
\]
The averaged POVM is covariant and has the same Bayes risk because the prior
and cost are left invariant. The covariant Naimark--imprimitivity theorem
\cite{holevo1982} then provides an isometry into
$L^2(\SO)\otimes\cK$ that intertwines $U$ with the left regular
representation and turns the measurement into the canonical position
measurement. A mixed state may first be purified with a rotation-trivial
ancilla; this only enlarges the multiplicity spaces. Thus the mean Casimir and
the numbers $v_j$ are preserved.

Use the normalized Peter--Weyl identification
\[
  L^2(\SO)\cong\bigoplus_{j\geq0}V_j\otimes V_j^*,
  \qquad
  f_{mn}^{(j)}(g)=\sqrt{d_j}\,[U_j(g)]_{mn},
  \qquad d_j=2j+1.
\]
Let $X$ be multiplication by the spin-one character $\chi_1$. The
Clebsch--Gordan product formula shows that $X$ has a block from input spin $k$
to output spin $j$ only when
\[
  j\in\{k-1,k,k+1\},
\]
with the $j=k=0$ diagonal absent. Let
\[
  C_{jk}:V_j\longrightarrow V_1\otimes V_k
\]
be the Clebsch--Gordan isometry in the displayed matrix-coefficient convention.
After identifying $V_k\otimes V_k^*$ with the Hilbert--Schmidt operators on
$V_k$, the product formula gives the block
\begin{equation}
  T_{jk}(Y)
  =\sqrt{\frac{d_k}{d_j}}\,
    C_{jk}^*(\1_3\otimes Y)C_{jk}.
  \label{eq:cg-block}
\end{equation}
The map $\Phi(Y)=C_{jk}^*(\1_3\otimes Y)C_{jk}$ is unital and completely
positive. Schur's lemma gives
\[
  \Tr_{V_1}(C_{jk}C_{jk}^*)=\frac{d_j}{d_k}\1_{V_k},
  \qquad
  \Tr\Phi(Y)=\frac{d_j}{d_k}\Tr Y.
\]
Kadison's inequality therefore implies
\[
  \lVert\Phi(Y)\rVert_2^2
  \leq\Tr\Phi(Y^*Y)
  =\frac{d_j}{d_k}\lVert Y\rVert_2^2,
\]
and \cref{eq:cg-block} yields $\lVert T_{jk}\rVert_{2\to2}\leq1$.
Tensoring with the identity on any rotation-trivial multiplicity space does
not change this norm.

Decompose the dilated vector as $\Psi=\sum_j\Psi_j$ with
$\lVert\Psi_j\rVert=v_j$, and let $X_{jk}=\Pi_jX\Pi_k$. The block contraction
and Cauchy--Schwarz give
\[
  |\langle\Psi_j,X_{jk}\Psi_k\rangle|\leq v_jv_k.
\]
For this covariantized fixed POVM, covariance lets us set the true orientation
to the identity. Since $\chi_1(g^{-1})=\chi_1(g)$, the character score is
\[
  S
  =\int_{\SO}\chi_1(\hat g^{-1})
      \langle\psi,\widetilde{\mathsf M}(d\hat g)\psi\rangle
  =\langle\Psi,X\Psi\rangle.
\]
Consequently
\begin{equation}
  S=\sum_{j,k}\langle\Psi_j,X_{jk}\Psi_k\rangle
  \leq\sum_{j\geq1}v_j^2+2\sum_{j\geq0}v_jv_{j+1}
  =v^{\mathsf T}\mathsf Fv,
  \label{eq:fusion-score}
\end{equation}
where
\[
  \mathsf F_{00}=0,
  \qquad \mathsf F_{jj}=1\ (j\geq1),
  \qquad \mathsf F_{j,j+1}=\mathsf F_{j+1,j}=1,
\]
and all other entries vanish.

\subsection*{Dirichlet and Hardy estimates}

The unweighted Dirichlet sum converges for $v\in\ell^2$, while finite
$\overline C$ ensures convergence of the weighted sums below. Taking the
finite identity to the limit gives
\begin{equation}
  3-v^{\mathsf T}\mathsf Fv
  =\sum_{j\geq0}(v_j-v_{j+1})^2+2v_0^2.
  \label{eq:dirichlet}
\end{equation}
Moreover, finite $\overline C$ implies
$(N+1)v_{N+1}^2\to0$. Hence the finite telescoping formula
\[
  \sum_{j=0}^{N}(j+1)(v_j^2-v_{j+1}^2)
  =\sum_{j=0}^{N}v_j^2-(N+1)v_{N+1}^2
\]
converges to the identity
\[
  1=\sum_{j\geq0}(j+1)(v_j^2-v_{j+1}^2)
\]
This identity and Cauchy--Schwarz give
\begin{align*}
  1
  &\leq
  \left[\sum_{j\geq0}(v_j-v_{j+1})^2\right]^{1/2}
  \left[\sum_{j\geq0}(j+1)^2(v_j+v_{j+1})^2\right]^{1/2},\\
  \sum_{j\geq0}(j+1)^2(v_j+v_{j+1})^2
  &\leq 4\overline C+2.
\end{align*}
The second line follows from $(x+y)^2\leq2(x^2+y^2)$:
\[
  2\sum_{j\geq0}(j+1)^2(v_j^2+v_{j+1}^2)
  =4\overline C+2.
\]
Hence
\begin{equation}
  \sum_{j\geq0}(v_j-v_{j+1})^2
  \geq\frac{1}{4\overline C+2}.
  \label{eq:hardy}
\end{equation}
Finally, \cref{eq:chordal-cost} implies
$R(\rho,\mathsf M)=(3-S)/4$. Combining
\cref{eq:fusion-score,eq:dirichlet,eq:hardy} proves
\[
  R(\rho,\mathsf M)\geq\frac{1}{16\overline C+8}.
\]

For strict spin support $j\leq J$, diagonalizing the finite tridiagonal matrix
$\mathsf F$ with the eigenvector
$v_n\propto\sin[(n+\tfrac12)q]$ and
$q=2\pi/(2J+3)$ gives
\[
  \lambda_{\max}(\mathsf F)=1+2\cos\!\left(\frac{2\pi}{2J+3}\right),
  \qquad
  \Rref\geq\sin^2\!\left(\frac{\pi}{2J+3}\right).
\]

\section{Reproducibility and Claim Boundary}
\label{app:reproducibility}

The theorem and model calculations are implemented without external numerical
data. From the repository root, run
\begin{verbatim}
PYTHONPATH=. python -m pytest -q \
  tests/test_locality_reference_leakage.py \
  tests/test_global_so3_reference_risk.py \
  tests/test_spacelike_replication_manuscript.py
PYTHONPATH=. python -m qgtoy locality-reference-leakage
\end{verbatim}
The focused tests check the compression identity, robust constants, risk
inversion, parameter consistency, finite Dicke-basis realizations, and the
certificate claim boundary.

\bibliographystyle{unsrt}
\bibliography{references}

\end{document}
