\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage[T1]{fontenc}

\newtheorem{theorem}{Theorem}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}[theorem]{Remark}

\newcommand{\MN}{M_N}
\newcommand{\CN}{\mathbb{C}^N}
\newcommand{\Sh}{\mathrm{Sh}}
\newcommand{\Resp}{\mathrm{Resp}}
\newcommand{\Tr}{\mathrm{Tr}}
\newcommand{\id}{\mathrm{id}}

\title{Screen Shadows Do Not Determine Finite Observer Algebras}
\author{Brian Fioca}
\date{Draft technical note, June 2026}

\begin{document}
\maketitle

\begin{abstract}
We study finite static-patch-inspired observer models with matrix observer
algebra $\MN$, diagonal screen algebra $\CN$, and a dephased abelian control.
For a declared class of screen-visible diagnostics, the quantum model and the
abelian control have identical screen shadows, while intrinsic operator
response distinguishes their observer algebras. We record three stabilizing
structures: a cutoff-compatible strong-continuity gate excluding
instantaneous dephasing, a finite-to-Type-$\mathrm{II}_1$ scaffold via cofinal
matrix inclusions, and a consecutive-cutoff embedding audit comparing a
trace-filled UCP baseline with harmonic, heat-kernel, and
Berezin-Toeplitz-inspired refinements. Finally, we state a conditional
continuum-lift obstruction: any proposed static-patch or dS/CFT dictionary
that factors only through convergent screen shadows is incomplete whenever an
operator-response witness persists.

This is a finite benchmark note with reproducible code. It is not a continuum
de Sitter theorem, not a dS/CFT construction, and not a proof of ER=EPR in de
Sitter.
\end{abstract}

\section{Claim Boundary}

The finite claim is:
\[
\text{screen-visible finite diagnostics do not determine the observer algebra;}
\]
\[
\text{operator response, continuity, and embedding data are needed.}
\]
The standard background includes finite matrix algebras, diagonal
subalgebras, Schur channels, finite stabilizer/OA-QEC-style diagnostics, UHF
inclusions, and operator-algebraic reconstruction language. The contribution
is not a new definition of holography or algebraic ER=EPR. It is a linked
finite benchmark and obstruction ledger designed to make one common pitfall
executable: a screen-only diagnostic can miss the noncommutative observer
algebra.

We separate the results into four claim categories.
\begin{itemize}[leftmargin=2em]
\item \textbf{Exact finite theorem:} a statement about the finite objects
defined here.
\item \textbf{Bounded certificate:} a checked finite family or
implementation-backed audit.
\item \textbf{Conditional lift theorem schema:} a theorem form that becomes
continuum-relevant only if stated lift conditions are supplied.
\item \textbf{Continuum speculation:} motivation only; not claimed as proved
here.
\end{itemize}

\section{Finite Observer Model}

Fix a cutoff $L$ and write $N=N_L$. The finite benchmark compares
\[
  A_L=M_N,\qquad S_L\simeq \mathbb{C}^N,\qquad D_L\simeq \mathbb{C}^N.
\]
Here $A_L$ is the quantum observer algebra, $S_L$ is the diagonal screen
algebra, and $D_L$ is the dephased abelian control. Let
\[
  E_{\mathrm{diag}}:M_N\to\mathbb{C}^N
\]
be diagonal conditional expectation in a chosen screen basis.

\begin{definition}[Screen-shadow class]
For this benchmark, $\mathrm{Scr}_N$ is the declared class of finite
screen-visible diagnostics: functionals that factor through diagonal screen
data, such as the diagonal state $E_{\mathrm{diag}}(\rho)$, diagonal
correlators, finite horizon-overlap counts, and low-order screen-restricted
transfer records. The screen shadow is the tuple
\[
  \Sh_N(A,\rho,\Lambda)=
  \bigl(F(A,\rho,\Lambda)\bigr)_{F\in \mathrm{Scr}_N}.
\]
This is a finite benchmark definition, not a canonical continuum definition of
a gravitational screen.
\end{definition}

\begin{definition}[Operator response]
$\Resp_N$ is the finite operator-response tuple. It may query non-diagonal
algebraic structure, including off-diagonal matrix units, commutators,
operator or relative-entropy response under a channel, and generator
covariance under cutoff refinement maps.
\end{definition}

\section{Main Results}

\begin{theorem}[Screen-shadow collision]
For every $N\ge 2$, the quantum observer algebra $M_N$ and the dephased
abelian control $\mathbb{C}^N$ have identical $\Sh_N$ values for all declared
screen diagnostics in $\mathrm{Scr}_N$, but their observer algebras are not
isomorphic and are separated by $\Resp_N$.
\end{theorem}

\begin{proof}
The screen algebra in both cases is the same diagonal algebra
$\mathbb{C}^N$. Any screen-shadow diagnostic is, by definition, a function
only of the diagonal state, diagonal correlators, and declared
screen-restricted transfer records. Those data are matched after applying
$E_{\mathrm{diag}}$. However, $M_N$ is noncommutative for $N\ge2$, while
$\mathbb{C}^N$ is abelian. In $M_N$,
\[
  [e_{12},e_{21}]=e_{11}-e_{22},
  \qquad \|[e_{12},e_{21}]\|=1,
\]
whereas all commutators vanish in $\mathbb{C}^N$. Thus the screen shadow
agrees while the operator response and observer algebra differ.
\end{proof}

\begin{proposition}[Finite dynamics preserve the split]
The finite static-patch regulator channels used in the benchmark are
CPTP/unital Schur or random-unitary channels that preserve the diagonal screen
shadow and act nontrivially on off-diagonal matrix units. Within the audited
finite regulator class, the screen-shadow collision and operator-response
separation persist.
\end{proposition}

\begin{proof}[Proof sketch]
A Schur multiplier with positive semidefinite kernel and unit diagonal is
completely positive and unital. Because it fixes diagonal matrix units, it
preserves the declared diagonal screen data. Finite phase-kick dilations give
random-unitary Schur channels by averaging phases over a finite environment.
The dephased control erases off-diagonal matrix units, while the quantum
regulator retains a controlled off-diagonal response inside the declared
finite family. The executable certificates audit positivity, trace
preservation, diagonal shadows, and response fields for the bounded regulator
stack.
\end{proof}

\begin{theorem}[Strong-continuity gate]
Let $\Lambda_L(\delta)=\exp(\delta G_L)$ be an identity-starting finite
semigroup and suppose $\|G_L\|\le \Gamma_L$. If
$\delta_L\Gamma_L\to0$, then
\[
  \|\Lambda_L(\delta_L)-I\|
  \le \exp(\delta_L\Gamma_L)-1\to0.
\]
Thus cutoff-compatible strong continuity rules out abrupt dephasing routes
without naming $M_N$, $\mathbb{C}^N$, or off-diagonal response as an axiom.
\end{theorem}

\begin{proof}
The exponential series gives
\[
  \exp(\delta G)-I=\sum_{r\ge1}\frac{(\delta G)^r}{r!}.
\]
Taking operator norm and using submultiplicativity yields
\[
  \|\exp(\delta G)-I\|
  \le \sum_{r\ge1}\frac{(\delta\|G\|)^r}{r!}
  \le \exp(\delta\Gamma)-1.
\]
The right-hand side tends to zero if $\delta_L\Gamma_L\to0$. Stationary
twirling or complete dephasing fails this gate because it jumps by norm one on
an off-diagonal matrix unit.
\end{proof}

\begin{proposition}[Cofinal Type-$\mathrm{II}_1$ scaffold]
Consecutive spherical cutoffs $N_L=(L+1)^2$ do not generally admit unital
full-matrix $*$-homomorphisms $M_{N_L}\to M_{N_{L+1}}$, because such a map
exists only when $N_L$ divides $N_{L+1}$. The cofinal factorial subsequence
\[
  L_k=(k+1)!-1,\qquad N_k=((k+1)!)^2
\]
does admit trace-preserving full-matrix inclusions by amplification. The
resulting quantum inductive limit is a UHF algebra whose tracial GNS closure
is the hyperfinite Type-$\mathrm{II}_1$ factor by the standard UHF
trace-closure theorem. The dephased diagonal sequence has the same levelwise
screen shadows and an abelian von Neumann limit.
\end{proposition}

\begin{proof}[Proof sketch]
A unital $*$-homomorphism $M_n\to M_m$ exists exactly when $m=nr$, in which
case it is conjugate to $A\mapsto A\otimes I_r$. Consecutive cutoffs
$(L+1)^2$ do not divide $(L+2)^2$ for $L\ge1$. On the factorial subsequence,
$N_{k+1}/N_k=(k+2)^2$, so the amplification map exists and preserves the
normalized trace. The diagonal control splits each screen atom into equal
subatoms, preserving diagonal trace data. The UHF/trace-closure statement is
standard operator-algebra input; the finite certificate does not claim this is
the canonical de Sitter static-patch algebra.
\end{proof}

\begin{proposition}[Consecutive cutoff refinement]
For $n\le m$, choose an isometry $V:\mathbb{C}^n\to\mathbb{C}^m$, let
$P=VV^*$ and $Q=I_m-P$, and define
\[
  \Phi(A)=VAV^*+\tau_n(A)Q.
\]
Then $\Phi:M_n\to M_m$ is unital, completely positive, and normalized-trace
preserving. For $A=e_{12}$ and $B=e_{21}$,
\[
  \|\Phi(AB)-\Phi(A)\Phi(B)\|=1/n.
\]
For static-patch dimensions $n=N_L=(L+1)^2$, this selected matrix-unit error
tends to zero. This is not a uniform approximate-homomorphism estimate on the
full unit ball.
\end{proposition}

\begin{proof}
Unitality follows from $\Phi(I_n)=P+Q=I_m$. Complete positivity follows
because $A\mapsto VAV^*$ is completely positive and
$A\mapsto\tau_n(A)Q$ is a positive functional times a positive operator. For
normalized traces,
\[
  \tau_m(\Phi(A))
  =\frac{\Tr_n(A)+\tau_n(A)(m-n)}{m}
  =\frac{\Tr_n(A)}{n}
  =\tau_n(A).
\]
For $A=e_{12}$ and $B=e_{21}$, one has $AB=e_{11}$ and
$\tau_n(A)=\tau_n(B)=0$, so
\[
  \Phi(AB)-\Phi(A)\Phi(B)=\frac{1}{n}Q,
\]
whose operator norm is $1/n$ when $m>n$.

For even $n<m$, let $A=\operatorname{diag}(I_{n/2},-I_{n/2})$. Then
$\tau_n(A)=0$ and $A^2=I_n$, so
\[
  \|\Phi(A^2)-\Phi(A)^2\|=\|I_m-VV^*\|=1.
\]
Thus the trace-filled map is not uniformly asymptotically multiplicative. Its
role is a selected-witness finite audit; a continuum construction requires
uniform control on a physically specified low-energy operator system or a
separately justified weaker topology.
\end{proof}

The executable audit also compares harmonic mode-label refinement,
heat-kernel Schur coarse graining after harmonic refinement, and a
Berezin-Toeplitz-inspired CP smoothing surrogate. These maps are not claimed
canonical. Their role is to show that the finite obstruction survives a small
family of physically motivated cutoff/coarse-graining audits rather than only
the factorial inclusion scaffold.

\begin{theorem}[Conditional continuum-lift obstruction]
Suppose two finite regulator sequences satisfy embedding or coarse-graining
maps between cutoffs, trace/state convergence, screen-shadow convergence,
strong-continuity or generator control, persistence of an operator-response
witness, and compatibility with the proposed limiting observer algebra. If
their limiting screen shadows agree but their limiting response witnesses
differ, then no dictionary that factors only through the limiting screen
shadow can determine the observer algebra.
\end{theorem}

\begin{proof}
Let $D$ be a dictionary that factors through the screen shadow. If two
sequences have identical limiting screen shadows, $D$ assigns them the same
output. If an operator-response witness has a nonzero limiting gap, the two
limiting observer-algebra candidates are distinguishable by operator response.
Therefore $D$ cannot be a complete invariant of the observer algebra.
\end{proof}

\section{Why This Is Not Trivialized by Diagonal Blindness}

The finite collision is intentionally sharp: screen shadows are diagonal or
screen-restricted, while the response witness is noncommutative. A skeptical
reader may say that diagonal probes cannot see off-diagonal structure. That is
true, and it is exactly the benchmark target.

The nontrivial package around that elementary observation is that the
screen-shadow class is explicit, the response witness is exact, KMS-like
conditions are separated from strong continuity, derived finite dynamics
preserve the split, exact cofinal inclusions and approximate consecutive UCP
refinements are audited, and the continuum-lift theorem states what any
screen-only dictionary would fail to see under explicit lift assumptions.

\section{Expert Question}

The most useful feedback question is:
\[
\text{Are the continuum-lift obstruction and cutoff embedding/coarse-graining}
\]
\[
\text{problem meaningful ways to formalize the limitation of diagonal/screen-shadow data?}
\]
If so, should the finite maps be organized through harmonic projection,
heat-kernel coarse graining, Berezin-Toeplitz-style quantization, a continuum
$L^2(S^2)$ projection, approximate UCP embeddings, or a modular/KMS-based
conditional expectation intrinsic to the observer algebra?

\section{Limitations}

This note does not prove a continuum de Sitter observer-algebra theorem, a
dS/CFT dictionary, literal ER=EPR in de Sitter, canonical status for the
factorial subsequence, or correctness of the trace-filled, harmonic,
heat-kernel, or Berezin-inspired cutoff maps as the physical static-patch
refinement. It also does not prove approximate-QEC stability in a Type-II or
Type-III limit.

\section{Related Work}

The operator-algebra QEC background includes Beny, Kempf, and Kribs,
\emph{Quantum Error Correction of Observables} (arXiv:0705.1574), Almheiri,
Dong, and Harlow, \emph{Bulk Locality and Quantum Error Correction in
AdS/CFT} (arXiv:1411.7041), and Harlow, \emph{The Ryu-Takayanagi Formula from
Quantum Error Correction} (arXiv:1607.03901). HaPPY-style holographic
tensor-network codes are due to Pastawski, Yoshida, Harlow, and Preskill,
\emph{Holographic quantum error-correcting codes} (arXiv:1503.06237). The
algebraic ER=EPR motivation is Engelhardt and Liu, \emph{Algebraic ER=EPR and
Complexity Transfer} (arXiv:2311.04281). The static-patch Type-II motivation
is Chandrasekaran, Longo, Penington, and Witten, \emph{An Algebra of
Observables for de Sitter Space} (arXiv:2206.10780). Approximate fuzzy-sphere
matrix convergence is motivated by Rieffel's work on matrix algebras
converging to the sphere in quantum Gromov-Hausdorff distance
(arXiv:math/0108005).

\section{Reproducibility}

The compact package is reproduced by:
\begin{verbatim}
PYTHONPATH=. python3 examples/reproduce_static_patch_package.py
PYTHONPATH=. python3 -m unittest \
  tests.test_static_patch_strong_continuity \
  tests.test_typeii_static_patch_limit \
  tests.test_inclusion_covariant_dynamics \
  tests.test_embedding_channels \
  tests.test_continuum_lift_obstruction
\end{verbatim}
The repository contains the frozen certificate artifacts and command index.
The tests make the finite theorem package auditable and regression-safe; they
do not replace the human proofs above or construct a continuum static patch.

\end{document}
